---
title: "02 — Jacobian Community Matrix"
author: "Nate Heili"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview

This script converts bootstrapped energy flux matrices (from `01_energy_flux.Rmd`)
into **Jacobian community matrices** following the approach of Bazin et al. 2025.
The Jacobian J describes the per-unit-biomass effect of each species on every other
species at equilibrium, and its eigenstructure characterizes community stability.

**Why basal biomass matters here:**

`fluxweb` calculates fluxes entirely from the consumer side — metabolic demand
drives flux, distributed by diet proportions. Basal resource biomass (B_resource)
does not enter the flux calculation. However, it enters the Jacobian in two places:

1. **Off-diagonal elements** — the per-biomass effect of a basal resource on a
   consumer, and vice versa:
   $$J_{consumer,resource} = \frac{ae \cdot F_{ij}}{B_{resource}}$$
   $$J_{resource,consumer} = \frac{-F_{ij}}{B_{resource}}$$

2. **Diagonal self-damping of basal resources** — unlike consumers, basal resources
   are not self-limited by metabolic loss. Their self-damping reflects intrinsic
   population growth rate at carrying capacity:
   $$J_{ii}^{basal} = -r_i(T)$$
   where $r_i(T)$ is temperature-dependent specific growth rate (day-1).

With near-zero placeholder biomasses, both terms become ecologically uninterpretable
(inflated interaction strengths, extreme diagonal dominance). This script replaces
placeholders with three empirically-grounded scenarios.

**Assumptions (Bazin et al. 2025 framework):**

1. **Steady-state** — the Jacobian is linearised around the equilibrium that the
   flux matrix represents. Each sampling date is treated as a snapshot of a local
   quasi-equilibrium.
2. **Type I linear functional responses** — per-capita effect of consumer on
   resource is constant regardless of density. Reasonable for generalist grazers
   in low-predation-risk systems but underestimates nonlinearity at high densities.

**Inputs:**
- `data/processed/results_flux.rds` — raw flux matrices per site × date × boot
- `data/processed/flux_boots.rds` — boot-level community summaries
- `data/community.rds` — biomass and metabolic rates per taxon × site × date × boot

**Outputs:**
- `data/processed/results_jacobian_[scenario].rds` — Jacobian matrices per scenario
- `data/processed/interactions_all.rds` — tidy interaction strength distributions
- `data/processed/jacobian_summary.rds` — scenario comparison summary

---

# Libraries

```{r libraries}
rm(list = ls())

library(tidyverse)
library(lubridate)
```

---

# Load inputs

```{r load-data}
results_flux <- readRDS("data/processed/results_flux.rds")
community    <- readRDS("data/community.rds") %>%
  mutate(date_id = as_date(date_id))

# Site order and colors (consistent with 01_energy_flux.Rmd)
site_order  <- c("st14", "oh2", "st7", "st9", "st6", "hver")
site_colors <- c(
  st14 = "#2C7BB6", oh2 = "#5AAE61", st7 = "#A6D96A",
  st9  = "#FDAE61", st6 = "#F46D43", hver = "#D73027"
)

basal_taxa <- c("amorphous_detritus", "animal", "cyanobacteria",
                "diatom", "filamentous", "green_algae", "plant_material")

sites  <- names(results_flux)
```

---

# Site temperatures

Used to parameterise temperature-dependent basal growth rates and GPP-based
biomass scenarios.

```{r site-temperatures}
k_B <- 8.617e-5  # Boltzmann constant (eV / K)

# Annual mean temperature per site from community data
site_temps <- community %>%
  filter(!taxon_id %in% basal_taxa) %>%
  group_by(site) %>%
  summarise(
    kelvin  = mean(kelvin,  na.rm = TRUE),
    celsius = mean(celcius, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(site = factor(site, levels = site_order, ordered = TRUE)) %>%
  arrange(site)

print(site_temps)
```

---

# Basal biomass scenarios

## Background

Basal resource standing stock biomass is not directly measured in the community
bootstrap data. We implement three scenarios reflecting contrasting hypotheses
about how warming shapes basal resource availability:

**Scenario 1 — Constant (null baseline):**
All basal resources assigned 1 g AFDM m-2 regardless of site or taxon. This
is the standard placeholder used by Gauzens et al. 2019 and Jochum et al. 2021,
and serves as a baseline against which the empirical scenarios are compared.

**Scenario 2 — GPP-scaled (warming increases basal biomass):**
Standing stock biomass estimated from gross primary production (GPP) following
O'Gorman et al. 2016 (Fig 4D):

$$\ln(GPP) = \ln(0.9857) + 0.0756 \cdot T_C$$

where GPP is in g C m-2 d-1. At equilibrium, standing stock equals production divided by specific loss rate B=GPP/r. The temperature dependence of standing stock therefore depends on the *difference* in activation energies between production (E_GPP = 0.54 eV; O'Gorman et al. 2016) and algal-specific growth rate (E_r), reflecting a scenario where algal production scales more steeply with temperature. This produces a positive relationship between standing stock and temperature (E_GPP > E_r). We acknowledge that empirical standing stocks at Hengill are more consistent with the temperature-invariant expectation (E_r = E_GPP) and treat this scenario as an upper bound.

Conversion chain:
- GPP (g O) to GPP (g C): divide by 2.67 (32 g O₂ / 12 g C)
- B_algae (g AFDM) = GPP (g C m-2 d-1) / r_algae (d-1) / 0.5 (C:AFDM ratio)
  where r_algae is the temperature-dependent specific growth rate (see below)

**Scenario 3 — Compositional shift: diatom decline / cyanobacteria proliferation:**
Warming shifts basal resource *composition* rather than total biomass. Diatoms —
the most palatable and nutritious resource — decline with warming as cyanobacteria proliferate and
competitively exclude them. Cyanobacteria are thought to be largely inedible
to stream invertebrates and occupy near-zero diet proportions in the gut content
analysis, meaning their high standing stock does not translate into consumer
energy gain. Total basal biomass *increases* under this scenario (more cyanobacteria
than diatoms lost), but effective consumer-resource coupling *weakens* because
the growing resource pool is increasingly inaccessible. This scenario is distinct
from Scenario 2 in that the mechanism operates through composition and palatability,
not production rate — and distinct from a simple grazing-pressure story because
the energy is unavailable to consumers rather than merely consumed faster.

## Temperature-dependent algal growth rate

Used to convert GPP flux to standing stock biomass in Scenario 2, and to
parameterise basal self-damping diagonals in all scenarios.

Specific growth rate follows the Metabolic Theory of Ecology (MTE):
$$r(T) = r_{ref} \cdot \exp\!\left(\frac{-E_r}{k_B} \left(\frac{1}{T} - \frac{1}{T_{ref}}\right)\right)$$
where:
- $r_{ref}$ = 0.15 day-1 at $T_{ref}$ = 293.15 K (20°C) — mid-range estimate
  for a mixed benthic algal assemblage (Eppley 1972; Friberg et al. 2009)
- $E_r$ = 0.32 - activation energy from MTE: biomass accumulates with warming  
- $k_B$ = Boltzmann constant

**Assumption in Scenario 2**
Standing stock biomass is estimated from gross primary production (GPP) following
O'Gorman et al. 2016 (Fig. 4d):

```{r basal-growth-rate}
E_r   <- 0.32       # activation energy for algal GPP at Hengill (eV)
r_ref <- 0.15       # specific growth rate at 20°C (d-1) - can vary widely by taxa
T_ref <- 293.15     # reference temperature (K) from MTE

# Temperature-dependent specific growth rate
algal_growth_rate <- function(kelvin) {
  r_ref * exp((-E_r / k_B) * (1/kelvin - 1/T_ref))
}

# Compute for each site
site_temps <- site_temps %>%
  mutate(
    r_algae = algal_growth_rate(kelvin)
  )

# Growth rates 
site_temps %>% select(site, celsius, r_algae) %>% print()
```

## Compute scenario biomasses


```{r basal-biomass-scenarios}
# Scenario 1: Constant
B_constant <- 1.0  # g AFDM m-2, Gauzens et al. 2019 / Jochum et al. 2021

# Scenario 2: GPP-scaled biomass 2-10 g AFDM m-2
# Total algal biomass increases with warming following the shape of the
# O'Gorman et al. 2016 GPP regression, but constrained to an ecologically
# realistic range (2-10 g AFDM m-2) consistent with Hengill biofilm
# measurements (Friberg et al. 2009; Gudmundsdottir et al. 2011).
# GPP is used to set the relative scaling across sites rather than as
# a direct biomass estimate, avoiding unrealistic values from the
# GPP/r equilibrium derivation.

B_gpp_scaled <- site_temps %>%
  mutate(
    GPP_gC_per_d  = 0.9857 * exp(0.0756 * celsius),   # O'Gorman et al. 2016
    # Rescale GPP to 2-10 g AFDM m-2 range
    GPP_min       = min(GPP_gC_per_d),
    GPP_max       = max(GPP_gC_per_d),
    B_algal_total = 2 + (10 - 2) * (GPP_gC_per_d - GPP_min) / (GPP_max - GPP_min)
  )

# Scenario 3: Resource shift — diatom decline / cyanobacteria increase with warming
# Net effect: total basal biomass increases with warming, but effective
# consumer-resource coupling WEAKENS as the edible fraction shrinks

T_cold <- min(site_temps$celsius)
T_warm <- max(site_temps$celsius)

build_basal_cyano_shift <- function(celsius) {
  w <- (celsius - T_cold) / (T_warm - T_cold)
  
  # Total algal biomass increases with warming
  # bounded by Gudmundsdóttir et al. 2011; Friberg et 2009 estimates (2-10 g m-2)
  B_algal_total <- exp(log(2) + (log(10) - log(2)) * w)
  
  # Cyanobacteria fraction increases with warming (2-70%)
  frac_cyano      <- 0.02 + (0.70 - 0.02) * w
  frac_palatable  <- 1 - frac_cyano   # diatom + green + filamentous combined
  
  c(
    # Palatable taxa: declining share of increasing total
    # Diatoms dominate palatable fraction (60%), green and filamentous split remainder
    diatom             = B_algal_total * frac_palatable * 0.60,
    green_algae        = B_algal_total * frac_palatable * 0.20,
    filamentous        = B_algal_total * frac_palatable * 0.20,
    # Cyanobacteria: growing share of increasing total
    cyanobacteria      = B_algal_total * frac_cyano,
    # Non-algal resources: site-invariant
    amorphous_detritus = 1,
    plant_material     = 1,
    animal             = 1
  )
}


map_dfr(site_temps$site, function(s) {
  temp <- site_temps %>% filter(site == s) %>% pull(celsius)
  bv   <- build_basal_cyano_shift(temp)
  tibble(site = s, taxon_id = names(bv), B_gAFDM = as.numeric(bv))
}) %>%
  pivot_wider(names_from = taxon_id, values_from = B_gAFDM) %>%
  print()

# Build per-site lookup tables for each scenario
basal_biomass_scenarios <- list(
  
  # Scenario 1: uniform 1 g AFDM m-2 for all taxa and sites (null baseline)
  constant = map(sites, function(s) {
    c(diatom = B_constant, cyanobacteria = B_constant, green_algae = B_constant,
      filamentous = B_constant, amorphous_detritus = B_constant,
      plant_material = B_constant, animal = B_constant)
  }) %>% set_names(sites),
  
  # Scenario 2: total algal biomass increases 2-10 g m-2 following GPP shape
  # composition fixed (diatom dominated); detritus and animal site-invariant
  gpp_scaled = map(sites, function(s) {
    B_tot <- B_gpp_scaled %>% filter(site == s) %>% pull(B_algal_total)
    c(
      diatom             = B_tot * 0.60,
      green_algae        = B_tot * 0.20,
      filamentous        = B_tot * 0.20,
      cyanobacteria      = 1, # consistent with placeholder constant scenario
      amorphous_detritus = 1,
      plant_material     = 1,
      animal             = 1
    )
  }) %>% set_names(sites),
  
  # Scenario 3: taxon-specific compositional shift
  cyano_shift = map(sites, function(s) {
    temp <- site_temps %>% filter(site == s) %>% pull(celsius)
    build_basal_cyano_shift(temp)
  }) %>% set_names(sites)
)
```

# Visualize basal resource scenarios

```{r visualize scenarios}
# Build tidy dataframe of all three scenarios for visualization
scenario_all <- map_dfr(names(basal_biomass_scenarios), function(sc) {
  map_dfr(sites, function(s) {
    bv <- basal_biomass_scenarios[[sc]][[s]]
    tibble(
      scenario = sc,
      site     = s,
      taxon_id = names(bv),
      B_gAFDM  = as.numeric(bv)
    )
  })
}) %>%
  mutate(
    site     = factor(site, levels = site_order, ordered = TRUE),
    scenario = factor(scenario,
                      levels = c("constant", "gpp_scaled", "cyano_shift"),
                      labels = c("Scenario 1: Constant",
                                 "Scenario 2: GPP-scaled",
                                 "Scenario 3: Cyano shift")),
    taxon_id = factor(taxon_id,
                      levels = c("diatom", "green_algae", "filamentous",
                                 "cyanobacteria", "amorphous_detritus",
                                 "plant_material", "animal"))
  )

# Stacked bar faceted by scenario
ggplot(scenario_all,
       aes(x = site, y = B_gAFDM, fill = taxon_id)) +
  geom_col(colour = "white", linewidth = 0.3) +
  facet_wrap(~ scenario, ncol = 1) +
  scale_fill_manual(values = c(
    diatom             = "#4E9BB5",  # blue — most palatable
    green_algae        = "#7CB47E",  # green — palatable
    filamentous        = "#A8D08D",  # light green — palatable
    cyanobacteria      = "#C85250",  # red — inedible
    amorphous_detritus = "#8B6F47",  # brown — detritus
    plant_material     = "#B5A26E",  # tan — allochthonous
    animal             = "#D4A0A0"   # pink — meiofauna
  )) +
  labs(
    x    = "Site (cold to warm)",
    y    = expression("Basal biomass (g AFDM m"^{-2}*")"),
    fill = "Resource",
    title    = "Basal resource biomass across three scenarios",
  ) +
  theme_classic(base_size = 12) +
  theme(
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 11),
    legend.position  = "right"
  )

```

---

# Jacobian construction functions

## `flux_to_jacobian_empirical()`

This is an expanded version of the Bazin et al. formulation that properly handles
basal resources:

**Off-diagonal elements (i ≠ j):**
$$J_{ij} = \frac{ae_j \cdot F_{ij}}{B_j} - \frac{F_{ji}}{B_j}$$

This is identical to Bazin — we divide all off-diagonal elements by the prey's
biomass (B_j, column). The change from `01_energy_flux.Rmd` is that B_j for
basal taxa now comes from the scenario vector rather than near-zero placeholders.

**Diagonal — consumers:**
$$J_{ii}^{consumer} = -\frac{d_i \cdot B_i + \sum_j F_{ij}}{B_i} = -d_i - \frac{\sum_j F_{ij}}{B_i}$$
where $d_i$ is mass-specific metabolic loss (J day-1 g-1) and $F_{ij}$ is flux
out of species i to all consumers.

**Diagonal — basal resources:**
$$J_{ii}^{basal} = -r_i(T)$$
where $r_i(T)$ is the temperature-dependent specific algal growth rate (day-1).
For detritus: assigned -0.01 day-1 (slow decomposition / slow replenishment).
For animal diet item: assigned intermediate -0.05 day-1.

This replaces the metabolic loss diagonal used for consumers — basal resources
are self-limited by growth, not metabolic demand.

```{r fn-jacobian}
flux_to_jacobian_empirical <- function(flux_mat,
                                       biomass_vec,     # consumer biomasses (g m-2)
                                       loss_vec,        # mass-specific metabolic loss (J d-1 g-1)
                                       ae_vec,          # assimilation efficiencies
                                       basal_bio_vec,   # basal biomass (g AFDM m-2)
                                       r_basal,         # algal growth rate (day-1)
                                       r_detritus = 0.01,  # decomposition / replenishment rate
                                       r_animal   = 0.05   # meiofaunal growth rate
                                       ) {
  
  n        <- nrow(flux_mat)
  sp_order <- rownames(flux_mat)
  
  # Identify basal vs consumer indices
  is_basal    <- sp_order %in% basal_taxa
  is_algal    <- sp_order %in% c("diatom", "filamentous", "green_algae", "cyanobacteria")
  is_detritus <- sp_order %in% c("amorphous_detritus", "plant_material")
  is_animal   <- sp_order %in% c("animal")
  is_consumer <- !is_basal
  
  # Build full biomass vector: consumers from bootstrap, basal from scenario
  full_bio <- biomass_vec
  for (tax in names(basal_bio_vec)) {
    if (tax %in% sp_order) {
      full_bio[sp_order == tax] <- basal_bio_vec[tax]
    }
  }
  
  # Guard: check for zero/NA in full_bio
  if (any(is.na(full_bio)) || any(full_bio <= 0)) {
    problem <- sp_order[is.na(full_bio) | full_bio <= 0]
    warning(paste("Zero/NA biomass for:", paste(problem, collapse = ", ")))
    return(NULL)
  }
  
  # --- Off-diagonal: Bazin formulation ---
  # J[i,j] = ae_j * F[j,i] / B_j - F[i,j] / B_j
  # flux_mat is prey(row) × consumer(col): F[prey, consumer]
  # t(flux_mat) is consumer(row) × prey(col): which consumer eats which prey
  # ae applied column-wise (ae belongs to the prey being eaten)
  
  jacob <- t(flux_mat) * ae_vec - flux_mat   # prey × prey matrix of net flows
  jacob <- sweep(jacob, MARGIN = 2, full_bio, "/")  # divide columns by biomass
  
  # Diagonal
  for (i in seq_len(n)) {
    
    if (is_consumer[i]) {
      # Consumer diagonal: metabolic loss + predation loss, normalised by biomass
      metabolic_loss <- loss_vec[i]                            # already per gram
      predation_loss <- sum(flux_mat[i, ]) / full_bio[i]      # flux out / biomass
      jacob[i, i]    <- -(metabolic_loss + predation_loss)
      
    } else if (is_algal[i]) {
      # Algal diagonal: intrinsic growth rate (self-damping)
      # Negative sign: at carrying capacity, net growth rate → 0, self-damping = r
      jacob[i, i] <- -r_basal
      
    } else if (is_detritus[i]) {
      jacob[i, i] <- -r_detritus
      
    } else if (is_animal[i]) {
      jacob[i, i] <- -r_animal
    }
  }
  
  rownames(jacob) <- colnames(jacob) <- sp_order
  jacob
}
```

## `extract_interactions()`

Extracts off-diagonal Jacobian elements as a tidy tibble for distribution analysis.

```{r fn-interactions}
extract_interactions <- function(J, consumer_only = TRUE) {
  
  sp_order <- rownames(J)
  if (consumer_only) {
    sp_keep  <- sp_order[!sp_order %in% basal_taxa]
    J        <- J[sp_keep, sp_keep, drop = FALSE]
    sp_order <- sp_keep
  }
  
  expand.grid(recipient = sp_order, actor = sp_order,
              stringsAsFactors = FALSE) %>%
    as_tibble() %>%
    filter(recipient != actor) %>%
    mutate(
      interaction_strength = map2_dbl(recipient, actor, ~ J[.x, .y]),
      direction = case_when(
        interaction_strength > 0 ~ "positive",
        interaction_strength < 0 ~ "negative",
        TRUE                     ~ "zero"
      )
    ) %>%
    filter(interaction_strength != 0)
}
```

---

# Get community str estimates for each boot

This code takes a while to run. It essentially organizes the community structure data for Jacobian estimation (biomasses, metabolic losses) with the energy flux matrices for east site, date, and bootstrap.

```{r}
# Pre-compute all per-site x date x boot community data ONCE
# outside any scenario loop — this never changes across scenarios

community_prepped <- map(sites, function(s) {
  message("Prepping community: ", s)
  
  dates_s <- names(results_flux[[s]])
  
  map(dates_s, function(d) {
    date_val  <- as_date(d)
    flux_ex   <- results_flux[[s]][[d]]$flux_list[[1]]  # example matrix
    sp_order  <- rownames(flux_ex)
    consumers <- sp_order[!sp_order %in% basal_taxa]
    boot_ids  <- names(results_flux[[s]][[d]]$flux_list)
    
    # Basal AE: pulled once per site x date (invariant across boots)
    basal_ae <- community %>%
      filter(site == s, date_id == date_val, taxon_id %in% basal_taxa) %>%
      group_by(taxon_id) %>%
      summarise(ae_temp = mean(ae_temp, na.rm = TRUE), .groups = "drop")
    
    # Global minimums for placeholder logic: once per site x date
    global_min_bio  <- community %>%
      filter(site == s, date_id == date_val,
             !taxon_id %in% basal_taxa, bio_g_m > 0) %>%
      pull(bio_g_m) %>% min(na.rm = TRUE)
    
    global_min_loss <- community %>%
      filter(site == s, date_id == date_val,
             !taxon_id %in% basal_taxa, metrate_J_day_gram > 0) %>%
      pull(metrate_J_day_gram) %>% min(na.rm = TRUE)
    
    # Pull all boots at once for this site x date, then split
    comm_all <- community %>%
      filter(site == s, date_id == date_val,
             taxon_id %in% consumers,
             boot_id %in% boot_ids) %>%
      group_by(boot_id, taxon_id) %>%
      summarise(
        bio_g_m            = mean(bio_g_m,            na.rm = TRUE),
        metrate_J_day_gram = mean(metrate_J_day_gram,  na.rm = TRUE),
        ae_temp            = mean(ae_temp,             na.rm = TRUE),
        .groups = "drop"
      )
    
    # Build pre-assembled vectors per boot
    boot_vecs <- map(boot_ids, function(b) {
      
      comm_b <- comm_all %>%
        filter(boot_id == b) %>%
        arrange(match(taxon_id, sp_order))
      
      if (nrow(comm_b) == 0) return(NULL)
      
      bio_vec  <- setNames(rep(0, length(sp_order)), sp_order)
      loss_vec <- setNames(rep(0, length(sp_order)), sp_order)
      ae_vec   <- setNames(rep(0, length(sp_order)), sp_order)
      
      bio_vec[comm_b$taxon_id]  <- comm_b$bio_g_m
      loss_vec[comm_b$taxon_id] <- comm_b$metrate_J_day_gram
      ae_vec[comm_b$taxon_id]   <- comm_b$ae_temp
      
      # Basal AE from community.rds
      ae_vec[basal_ae$taxon_id] <- basal_ae$ae_temp
      
      # Placeholder for zero consumer biomass
      consumer_idx <- sp_order %in% consumers
      zero_idx     <- consumer_idx & bio_vec == 0
      if (any(zero_idx)) {
        bio_vec[zero_idx]  <- global_min_bio  * 0.01
        loss_vec[zero_idx] <- global_min_loss * 0.01
        ae_vec[zero_idx & ae_vec == 0] <- 0.906
      }
      
      list(bio_vec = bio_vec, loss_vec = loss_vec, ae_vec = ae_vec)
    }) %>% set_names(boot_ids)
    
    list(
      boot_vecs = boot_vecs,
      sp_order  = sp_order
    )
    
  }) %>% set_names(dates_s)
  
}) %>% set_names(sites)

message("Community pre-computation complete")
```

# Build Jacobians across all scenarios

```{r constant}
sc <- "constant"
message("Running scenario: ", sc)

results_jacobian_constant <- map(sites, function(s) {
  basal_bio <- basal_biomass_scenarios[[sc]][[s]]
  r_site    <- site_temps %>% filter(site == s) %>% pull(r_algae)
  
  map(names(results_flux[[s]]), function(d) {
    prepped  <- community_prepped[[s]][[d]]
    flux_res <- results_flux[[s]][[d]]
    
    jacob_list <- map(names(flux_res$flux_list), function(b) {
      vecs <- prepped$boot_vecs[[b]]
      if (is.null(vecs)) return(NULL)
      
      tryCatch(
        flux_to_jacobian_empirical(
          flux_mat      = flux_res$flux_list[[b]],
          biomass_vec   = vecs$bio_vec,
          loss_vec      = vecs$loss_vec,
          ae_vec        = vecs$ae_vec,
          basal_bio_vec = basal_bio,
          r_basal       = r_site
        ),
        error = function(e) NULL
      )
    }) %>% set_names(names(flux_res$flux_list))
    
    Filter(Negate(is.null), jacob_list)
    
  }) %>% set_names(names(results_flux[[s]]))
  
}) %>% set_names(sites)

saveRDS(results_jacobian_constant, "data/processed/results_jacobian_constant.rds")
```





```{r main-loop, cache=TRUE}
sc <- "gpp_scaled"
message("Running scenario: ", sc)

results_jacobian_GPPscaled <- map(sites, function(s) {
  basal_bio <- basal_biomass_scenarios[[sc]][[s]]
  r_site    <- site_temps %>% filter(site == s) %>% pull(r_algae)
  
  map(names(results_flux[[s]]), function(d) {
    prepped  <- community_prepped[[s]][[d]]
    flux_res <- results_flux[[s]][[d]]
    
    jacob_list <- map(names(flux_res$flux_list), function(b) {
      vecs <- prepped$boot_vecs[[b]]
      if (is.null(vecs)) return(NULL)
      
      tryCatch(
        flux_to_jacobian_empirical(
          flux_mat      = flux_res$flux_list[[b]],
          biomass_vec   = vecs$bio_vec,
          loss_vec      = vecs$loss_vec,
          ae_vec        = vecs$ae_vec,
          basal_bio_vec = basal_bio,
          r_basal       = r_site
        ),
        error = function(e) NULL
      )
    }) %>% set_names(names(flux_res$flux_list))
    
    Filter(Negate(is.null), jacob_list)
    
  }) %>% set_names(names(results_flux[[s]]))
  
}) %>% set_names(sites)

saveRDS(results_jacobian_GPPscaled, "data/processed/results_jacobian_GPPscaled.rds")
```

```{r}
sc <- "cyano_shift"
message("Running scenario: ", sc)

results_jacobian_cyano_shift <- map(sites, function(s) {
  basal_bio <- basal_biomass_scenarios[[sc]][[s]]
  r_site    <- site_temps %>% filter(site == s) %>% pull(r_algae)
  
  map(names(results_flux[[s]]), function(d) {
    prepped  <- community_prepped[[s]][[d]]
    flux_res <- results_flux[[s]][[d]]
    
    jacob_list <- map(names(flux_res$flux_list), function(b) {
      vecs <- prepped$boot_vecs[[b]]
      if (is.null(vecs)) return(NULL)
      
      tryCatch(
        flux_to_jacobian_empirical(
          flux_mat      = flux_res$flux_list[[b]],
          biomass_vec   = vecs$bio_vec,
          loss_vec      = vecs$loss_vec,
          ae_vec        = vecs$ae_vec,
          basal_bio_vec = basal_bio,
          r_basal       = r_site
        ),
        error = function(e) NULL
      )
    }) %>% set_names(names(flux_res$flux_list))
    
    Filter(Negate(is.null), jacob_list)
    
  }) %>% set_names(names(results_flux[[s]]))
  
}) %>% set_names(sites)

saveRDS(results_jacobian_cyano_shift, "data/processed/results_jacobian_cyano_shift.rds")

```

---

# Extract interaction strength distributions

For each scenario, extract off-diagonal Jacobian elements and summarize (mean +/- 95% bootstrap interval). This links energy fluxes to the distribution of interaction strengths.

```{r extract-interactions}

summarise_jacobian_boots <- function(jacobian_scenario) {
  
  map(names(jacobian_scenario), function(s) {
    message("Summarising: ", s)
    
    map(names(jacobian_scenario[[s]]), function(d) {
      
      boot_mats <- jacobian_scenario[[s]][[d]]
      boot_mats <- Filter(Negate(is.null), boot_mats)
      if (length(boot_mats) == 0) return(NULL)
      
      # Stack all boot matrices into a 3D array: [row, col, boot]
      sp_order <- rownames(boot_mats[[1]])
      n        <- length(sp_order)
      n_boots  <- length(boot_mats)
      
      arr <- array(
        unlist(boot_mats),
        dim = c(n, n, n_boots),
        dimnames = list(sp_order, sp_order, names(boot_mats))
      )
      
      # Element-wise mean and 95% CI across boots (3rd dimension)
      J_mean <- apply(arr, c(1, 2), mean,            na.rm = TRUE)
      J_lo   <- apply(arr, c(1, 2), quantile, 0.025, na.rm = TRUE)
      J_hi   <- apply(arr, c(1, 2), quantile, 0.975, na.rm = TRUE)
      
      list(
        mean    = J_mean,
        lo      = J_lo,
        hi      = J_hi,
        n_boots = n_boots
      )
      
    }) %>% set_names(names(jacobian_scenario[[s]]))
    
  }) %>% set_names(names(jacobian_scenario))
}

# Run for each scenario
message("Summarising constant...")
jacobian_summary_constant    <- summarise_jacobian_boots(results_jacobian_constant)

message("Summarising GPP-scaled...")
jacobian_summary_gpp         <- summarise_jacobian_boots(results_jacobian_GPPscaled)

message("Summarising cyano shift...")
jacobian_summary_cyano       <- summarise_jacobian_boots(results_jacobian_cyano_shift)



# Fix alignment — pivot all three matrices identically
mat_to_long <- function(mat) {
  as.data.frame(mat) %>%
    rownames_to_column("recipient") %>%
    pivot_longer(-recipient, names_to = "actor", values_to = "value")
}


# Make a df of interaction strengths from the Jacobian
interactions_all <- map_dfr(
  list(constant    = jacobian_summary_constant,
       gpp_scaled  = jacobian_summary_gpp,
       cyano_shift = jacobian_summary_cyano),
  function(sc_summary) {
    map_dfr(sites, function(s) {
      map_dfr(names(sc_summary[[s]]), function(d) {
        
        res <- sc_summary[[s]][[d]]
        if (is.null(res)) return(NULL)
        
        # Pivot all three matrices the same way then join
        df_mean <- mat_to_long(res$mean) %>% rename(mean_strength = value)
        df_lo   <- mat_to_long(res$lo)   %>% rename(lo_strength   = value)
        df_hi   <- mat_to_long(res$hi)   %>% rename(hi_strength   = value)
        
        df_mean %>%
          left_join(df_lo, by = c("recipient", "actor")) %>%
          left_join(df_hi, by = c("recipient", "actor")) %>%
          filter(recipient != actor, mean_strength != 0) %>%
          mutate(
            site               = s,
            date               = as_date(d),
            log10_abs_mean     = log10(abs(mean_strength)),
            log10_abs_lo       = log10(abs(lo_strength)),
            log10_abs_hi       = log10(abs(hi_strength)),
            direction          = if_else(mean_strength > 0, "positive", "negative"),
            recipient_is_basal = recipient %in% basal_taxa,
            actor_is_basal     = actor     %in% basal_taxa,
            interaction_type   = case_when(
              !recipient_is_basal & actor_is_basal  ~ "consumer_resource",
              recipient_is_basal  & !actor_is_basal ~ "resource_consumer",
              !recipient_is_basal & !actor_is_basal ~ "consumer_consumer",
              TRUE                                  ~ "resource_resource"
            )
          ) %>%
          select(-recipient_is_basal, -actor_is_basal)
      })
    })
  },
  .id = "scenario"
) %>%
  mutate(
    site     = factor(site,     levels = site_order, ordered = TRUE),
    scenario = factor(scenario, levels = c("constant", "gpp_scaled", "cyano_shift"))
  )

# Remove placeholder taxa — that have a near-zero biomass value and inflate J
interactions_real <- interactions_all %>%
  filter(round(lo_strength, 8) != round(hi_strength, 8)) %>%
  mutate(julian = yday(date))
```

---

# Visualise interaction strength distributions

## Consumer-resource coupling |J[consumer,resource]|
We quantified consumer-resource coupling as the off-diagonal Jacobian elements J[consumer, resource] = ae × F_rc / B_resource, representing the per-unit-biomass effect of each basal resource on each consumer's population growth rate at equilibrium. These elements integrate empirically-estimated energy fluxes (F_rc; J m-2 day-1) with temperature-adjusted assimilation efficiencies (ae) and scenario-specific basal resource biomass (B_resource; g AFDM m-2), yielding interaction strengths in units of day-1. For each site and sampling date, we summarized the distribution of J[consumer, resource] across all realized consumer-resource pairs, excluding taxa with near-zero placeholder biomass that produced uninformative near-zero interactions. We report the median and interquartile range of |J[c,r]| as a measure of community-wide coupling strength, where higher values indicate stronger per-biomass consumer dependence on basal resources and greater potential for resource fluctuations to propagate into consumer dynamics.

```{r |J(c,r)|}
# Quick boxplot visual of J[c,r] across gradient, faceted by scenario
interactions_real %>%
  filter(interaction_type == "consumer_resource",
         is.finite(log10_abs_mean)) %>%
  ggplot(aes(x = site, y = log10_abs_mean, colour = site)) +
  geom_boxplot() +
  facet_wrap(~ scenario) +
  scale_colour_manual(values = site_colors) +
  labs(x = "Site (cold to warm)",
       y = expression(log[10]*"|J[consumer,resource]|",
       title = "Consumer-resource coupling across the thermal gradient")) +
  theme_classic() +
  theme(legend.position = "none")

# Interactions through time
interactions_real %>%
  filter(interaction_type == "consumer_resource",
         scenario == "constant",
         is.finite(log10_abs_mean)) %>%
  group_by(site, date) %>%
  summarise(
    median_coupling = median(log10_abs_mean),
    lo_coupling     = quantile(log10_abs_mean, 0.25),
    hi_coupling     = quantile(log10_abs_mean, 0.75),
    .groups = "drop"
  ) %>%
  mutate(
    doy  = yday(date),
    site = factor(site, levels = site_order, ordered = TRUE)
  ) %>%
  ggplot(aes(x = doy, y = median_coupling,
             colour = site, fill = site)) +
  geom_ribbon(aes(ymin = lo_coupling, ymax = hi_coupling),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_colour_manual(values = site_colors, name = "Site") +
  scale_fill_manual(values = site_colors,   name = "Site") +
  scale_x_continuous(
    breaks = yday(as.Date(paste0("2011-", 1:12, "-01"))),
    labels = month.abb,
    limits = c(1, 365)
  ) +
  labs(
    x        = NULL,
    y        = expression("log|J[consumer,resource]|"),
    title    = "Seasonal dynamics of consumer-resource coupling",
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "right",
    axis.text.x     = element_text(size = 9)
  ) +
  facet_wrap(~site)

# Check how scenarios influence interaction strengths
interactions_real %>%
  filter(interaction_type == "consumer_resource",
         is.finite(log10_abs_mean)) %>%
  group_by(scenario, site) %>%
  summarise(
    median_coupling = median(log10_abs_mean),
    lo_coupling     = quantile(log10_abs_mean, 0.25),
    hi_coupling     = quantile(log10_abs_mean, 0.75),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = site, y = median_coupling,
             colour = scenario, group = scenario)) +
  geom_pointrange(
    aes(ymin = lo_coupling, ymax = hi_coupling),
    position = position_dodge(width = 0.4),
    size = 0.6
  ) +
  scale_colour_manual(
    values = c(constant = "grey40", gpp_scaled = "#E69F00", cyano_shift = "#C85250"),
    labels = c("Constant", "GPP-scaled", "Cyano shift")
  ) +
  labs(
    x      = "Site (cold → warm)",
    y      = expression("Median log"[10]*"|J[consumer,resource]|"),
    colour = "Basal biomass scenario",
    title  = "Scenario sensitivity: consumer-resource coupling",
    subtitle = "Median +/- 95% bootstrap interval"
  ) +
  theme_classic(base_size = 12)

```
# Results: Interaction strengths from the Jacobian matrix
Consumer-resource interactions increased generally with warming, with the highest values in hver. Interstingly, there was lower CR coupling in st6 and st9 (17.6 and 11.1 C, respectively) relative to oh2 and st7 (5.64 and 5.98 C, respectively), a result that follows the energy flux patterning. oh2 and st7 have diverse communities (N = 31) with fairly high biomasses, which increase their per-capita effect on resources. 

Scenario specific changes to basal biomasses only influenced interactions in the warmer streams (hver, st6, st9). Increasing basal biomasses slightly lowered interaction strengths, or how tightly consumers are coupled to their resources, but if compisition shifts towards cyanobacteria, interactions go up (based on GCA analysis). This shows that there are potential for compensatory dynamics from increased primary production on the stability in warmed systems; consumers can offset some of their energetic demand if food resources are available to them. I think these scenarios will be more interesting as I develop ways to incorporate real basal resource biomass into the energy flux analysis and Jacobian.

Seasonal dynamics of CR interactions were most dynamic in the warmest stream - hver - where interaction strengths went up during hotter summer months. This results is driven by increases in consumer biomasses - notably Naidid + Cricotopus - increasing their overall energetic needs relative to the set basal resource stock. This may be an artifcat of the way it is modeled here where basal resources have no temporal dynamics, which is clearly not the case as primary production follows a somewhat similar patterning to consumers, peaking in the summer. Incorportaing temproal patterning may offset these coupling effects. Interestingly cold sites (st14) show the opposite: coupling is generally weak and relatively flat seasonally, potentially from low-biomass slow-turnover communities.

## Consumer-resource interactions |J(resource,consumer)| - Predation pressure

Ecologically these elements of the Jacobian represent how much a consumer supresses its resource per unit consumer biomass per day. Higher values mean the consumers has stronger top-down efect on the resource. The values are assymetric - a consumer can be weakly stimulated by a resource (low J[c,r]) but still exter strong suppression on it (high J[r,c]) if consumer biomass is high relative to flux.

```{r |J(r,c)|}
# Grazing pressure: J[resource, consumer]
interactions_real %>%
  filter(interaction_type == "resource_consumer",
         is.finite(log10_abs_mean)) %>%
  group_by(site) %>%
  summarise(
    median_grazing = median(log10_abs_mean),
    lo_grazing     = quantile(log10_abs_mean, 0.25),
    hi_grazing     = quantile(log10_abs_mean, 0.75),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = site, y = median_grazing,
             colour = site, group = site)) +
  geom_pointrange(
    aes(ymin = lo_grazing, ymax = hi_grazing),
    position = position_dodge(width = 0.4),
    size = 0.6
  ) +
  scale_colour_manual(values = site_colors
  ) +
  labs(
    x        = "Site (cold to warm)",
    y        = expression("Median log|J[resource,consumer]|"),
    colour   = "Site",
    title    = "Grazing pressure across the thermal gradient",
    subtitle = "Per-unit-consumer-biomass suppression of basal resources"
  ) +
  theme_classic(base_size = 12)

interactions_real %>%
  filter(interaction_type == "resource_consumer",
         scenario == "constant",
         is.finite(log10_abs_mean)) %>%
  group_by(site, date) %>%
  summarise(
    median_coupling = median(log10_abs_mean),
    lo_coupling     = quantile(log10_abs_mean, 0.25),
    hi_coupling     = quantile(log10_abs_mean, 0.75),
    .groups = "drop"
  ) %>%
  mutate(
    doy  = yday(date),
    site = factor(site, levels = site_order, ordered = TRUE)
  ) %>%
  ggplot(aes(x = doy, y = median_coupling,
             colour = site, fill = site)) +
  geom_ribbon(aes(ymin = lo_coupling, ymax = hi_coupling),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_colour_manual(values = site_colors, name = "Site") +
  scale_fill_manual(values = site_colors,   name = "Site") +
  scale_x_continuous(
    breaks = yday(as.Date(paste0("2011-", 1:12, "-01"))),
    labels = month.abb,
    limits = c(1, 365)
  ) +
  labs(
    x        = NULL,
    y        = expression("log|J[resource,consumer]|"),
    title    = "Seasonal dynamics of predation pressure",
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "right",
    axis.text.x     = element_text(size = 9)
  )

```

---

# Save outputs

```{r save-outputs}
# I use the constant scenario from above for stability given limited changes to Jacobian matrices no matter basal biomass
results_jacobian <- results_jacobian_constant

dir.create("data/processed", showWarnings = FALSE, recursive = TRUE)

# Full nested Jacobian list (input for 03_stability.Rmd)
saveRDS(results_jacobian, "data/processed/results_jacobian.rds")

```
