---
title: "eflux_bio_bootstrap"
author: "Nate Heili"
date: "2026-03-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Using energetic food web models to assess ecological stability

*"Food webs and their models are at best caricatures of reality, and thus have both the truth and the falsity of caricature."* May 1973

Ecologists have long sought to understand how ecosystems respond to disturbances. Theory and empirical work suggest that **interaction strengths** within complex communities are key determinants of how a system responds to perturbation. Specifically, communities with a high frequency of weak interactions can dampen the effects of species loss following a disturbance. One way to quantify interaction strengths is by estimating energy fluxes through the food web network, which can be used to construct a Jacobian matrix describing the per-capita effect of each species on every other. Eigenanalysis of the Jacobian then characterizes the dynamics of pertubations and global stability, with the real part of the dominant eigenvalue indicating whether the community will return to quilibrium following a disturbance.

In this tutorial, I use empirical food web data from a broad stream thermal gradient (Junker et al. 2024) to construct Jacobian matrices of interaction strengths using a bioenergetic modelling approach in the *fluxweb* package (Moore and de Ruiter 2012; Gauzens et al. 2019). To propogate uncertainty in population biomass estimates, I iterate this process across bootstrapped communities, yielding distributions of energy fluxes and stability estimates rather than single point estimates.  

```{r}
# Clean environemnt 
#rm(list = ls())

# load packages
library(fluxweb)
library(tidyverse)
library(readr)
library(ggridges)

```

# Load data
Data for the bioenergetic model (from Junker et al.) are stored on my GitHub page *https://github.com/nheili42/stability-tutorial*. The metadata file explaining the contents of each data source is on Github. I load the cleaned community dataframe and the associated diet matrix. Code to reproduce data wrangling is found on Github. 

The community df includes bootstrap estimates of pop biomass for each site on a given date, with the associated temp-dependent metabolic rate (Jochum et al. 2017; Brown et al. 2004) and assimilation efficiency (Lang et al. 2017). The diet matrix includes seasonal and annual diet proportions, summarized from a Bayesian model. This matrix will need to be cleaned up to match the community df inputs once we use the fluxing function from fluxwebs.

```{r}
community <- readRDS("data/community.rds")
diet <- readRDS("data/diet_matrices_clean.rds")
```

# Fluxing and stability
I estimate energy fluxes across bootstrap estimates, giving a range of plausible fluxes based on variation in population biomasses. Diet matrix proportions were derived from gut content analysis and therefore represent realized feeding proportions; consequently bioms.prefs = FALSE was used so that observed proportions were not overridden by biomass-weighting. I decided to set basal resource biomass to 0 mg/m2 - I am exploring a way to think about how we might quantify these biomass stocks for my PhD. Either way, we will be able to estimate the amount of each resource needed to support consumers, therefore be able to compare across streams. 

```{r load functions for energy fluxes}

# =============================================================================
# FUNCTION 1: prep_community()
# Filter community data by site x season, aggregate boots, fix zero biomass
# =============================================================================

prep_community <- function(community_df, site_name, season_name = NULL, diet_mat) {
  
  sp_order <- rownames(diet_mat)
  
  df <- community_df %>% filter(site == site_name)
  if (!is.null(season_name)) df <- df %>% filter(season == season_name)
  if (nrow(df) == 0) return(NULL)
  
  if (!all(sp_order %in% unique(df$taxon_id))) {
    missing <- setdiff(sp_order, unique(df$taxon_id))
    warning(paste("Missing taxa for", site_name,
                  ifelse(is.null(season_name), "annual", season_name), ":",
                  paste(missing, collapse = ", ")))
    return(NULL)
  }
  
  agg <- df %>%
    group_by(boot_id, taxon_id) %>%
    summarise(
      bio_g_m            = mean(bio_g_m,            na.rm = TRUE),
      metrate_J_day_gram = mean(metrate_J_day_gram,  na.rm = TRUE),
      ae_temp            = mean(ae_temp,             na.rm = TRUE),
      .groups = "drop"
    )
  
  # global fallback so always-zero taxa are also covered ---
  global_min_bio  <- agg %>% filter(bio_g_m > 0)            %>% pull(bio_g_m)            %>% min(na.rm = TRUE)
  global_min_loss <- agg %>% filter(metrate_J_day_gram > 0) %>% pull(metrate_J_day_gram) %>% min(na.rm = TRUE)
  
  # Per-taxon minimum where available, global minimum as backstop
  placeholders <- agg %>%
    filter(bio_g_m > 0 & metrate_J_day_gram > 0) %>%
    group_by(taxon_id) %>%
    summarise(
      ph_bio  = min(bio_g_m,            na.rm = TRUE) * 0.01,
      ph_loss = min(metrate_J_day_gram,  na.rm = TRUE) * 0.01,
      .groups = "drop"
    )
  
  # Full species list so always-zero taxa get the global fallback
  all_taxa <- tibble(taxon_id = sp_order) %>%
    left_join(placeholders, by = "taxon_id") %>%
    mutate(
      ph_bio  = coalesce(ph_bio,  global_min_bio  * 0.01),
      ph_loss = coalesce(ph_loss, global_min_loss * 0.01)
    )
  
  agg %>%
    left_join(all_taxa, by = "taxon_id") %>%
    mutate(
      bio_g_m            = case_when(
        is.na(bio_g_m)  | bio_g_m  == 0 ~ ph_bio,
        TRUE                             ~ bio_g_m),
      metrate_J_day_gram = case_when(
        is.na(metrate_J_day_gram) | metrate_J_day_gram == 0 ~ ph_loss,
        TRUE                                                  ~ metrate_J_day_gram)
    ) %>%
    select(-ph_bio, -ph_loss)
}

# =============================================================================
# FUNCTION 2: run_fluxes()
# Run fluxing() across all bootstraps for one site x temporal scale
# =============================================================================

run_fluxes <- function(community_agg, diet_mat) {
  
  sp_order <- rownames(diet_mat)
  boot_ids <- unique(community_agg$boot_id)
  
  flux_list <- lapply(boot_ids, function(b) {
    
    df_b <- community_agg %>%
      filter(boot_id == b) %>%
      arrange(match(taxon_id, sp_order))
    
    if (nrow(df_b) != nrow(diet_mat)) {
      warning(paste("Boot", b, "row mismatch — skipping"))
      return(NULL)
    }
    if (!all(df_b$taxon_id == sp_order)) {
      warning(paste("Boot", b, "order mismatch — skipping"))
      return(NULL)
    }
    
    tryCatch(
      fluxing(
        mat          = diet_mat,
        biomasses    = df_b$bio_g_m,
        losses       = df_b$metrate_J_day_gram,
        efficiencies = df_b$ae_temp,
        bioms.prefs  = FALSE,  # GCA diet proportions used as-is
        bioms.losses = TRUE,   # losses are per-gram, scaled internally
        ef.level     = "prey"  # ae is a property of the prey
      ),
      error = function(e) { warning(paste("Boot", b, ":", e$message)); NULL }
    )
  })
  
  names(flux_list) <- boot_ids
  Filter(Negate(is.null), flux_list)
}

```

# Jacobian matrices for stability
Once I derive energy fluxes, I convert them to Jacobian matrices which describe the local linear dynamics of the consumer community at equilibrium. Each element quantifies the per-capita effect of one taxon on another's instantaneous rate of change, with signs and magnitudes determined by the underlying trophic interaction structure and energetic constraints. From the eigenvalues of this community matrix, I extract two stability metrics: resilience, the rate at which the community returns to equilibrium following a small perturbation, and reactivity, the maximum rate at which perturbations amplify before that return begins. I follow the analytical approach described in Bazin et al., 2025.

```{r load functions for Jacobian}

# =============================================================================
# FUNCTION 3: flux_to_jacobian()
# Bazin et al. create.jacob formulation WITH self-damping diagonal
#
# Off-diagonal: identical to Bazin —
#   jacob = t(flux_mat) * ae - flux_mat
#   jacob = sweep(jacob, MARGIN=2, biomasses, '/')
#
# Diagonal (added here, absent in Bazin):
#   J[i,i] = -(total metabolic loss + total predation loss on i) / biomass[i]
#   = -(loss_per_gram[i] * biomass[i] + sum(flux_mat[i,])) / biomass[i]
#   = -loss_per_gram[i] - sum(flux_mat[i,]) / biomass[i]
# =============================================================================

flux_to_jacobian <- function(flux_mat, biomass_vec, loss_vec, ae_vec) {
  
  n        <- nrow(flux_mat)
  sp_order <- rownames(flux_mat)
  
  # Guard: check for zero/NA biomass before proceeding
  if (any(is.na(biomass_vec)) || any(biomass_vec == 0)) {
    problem_sp <- sp_order[is.na(biomass_vec) | biomass_vec == 0]
    warning(paste("Zero/NA biomass for:", paste(problem_sp, collapse = ", "),
                  "— Jacobian will have Inf. Returning NULL."))
    return(NULL)
  }
  
  # Off-diagonal: Bazin formulation
  # t(flux_mat) * ae_vec: gains (prey j -> predator i, scaled by ae of prey j)
  # - flux_mat: losses (flux out of i to all consumers)
  jacob <- t(flux_mat) * ae_vec - flux_mat
  
  # Divide each column j by biomass of species j
  jacob <- sweep(jacob, MARGIN = 2, biomass_vec, "/")
  
  # Self-damping diagonal (departure from Bazin)
  # Total loss on species i = metabolic loss + predation loss
  # Both divided by biomass[i] to get per-capita rate
  for (i in seq_len(n)) {
    if (biomass_vec[i] > 0) {
      metabolic_loss  <- loss_vec[i]                           # already per gram
      predation_loss  <- sum(flux_mat[i, ]) / biomass_vec[i]  # flux out / biomass
      jacob[i, i]     <- -(metabolic_loss + predation_loss)
    } else {
      jacob[i, i] <- 0
    }
  }
  
  rownames(jacob) <- colnames(jacob) <- sp_order
  jacob
}


# =============================================================================
# FUNCTION 4: jacobian_stability()
# Following the approach of Bazin et al. stability metrics
# Resilience  = max(Re(eigenvalues(J)))       positive = unstable
# Reactivity  = max(eigenvalues(H))           H = (J + J^T) / 2
# =============================================================================

jacobian_stability <- function(J) {
  
  eigs        <- eigen(J, only.values = TRUE)$values
  resilience  <- max(Re(eigs))
  
  J_hermitian <- (J + t(Conj(J))) / 2
  reactivity  <- max(Re(eigen(J_hermitian, only.values = TRUE)$values))
  
  tibble(
    resilience = resilience,    # λJmax: more positive = more unstable (long-term)
    reactivity = reactivity,    # λHmax: more positive = more reactive (short-term)
    is_stable  = resilience < 0
  )
}


# =============================================================================
# FUNCTION 5: extract_interaction_strengths()
# Off-diagonal Jacobian elements as tidy tibble
# =============================================================================

extract_interaction_strengths <- function(J) {
  
  sp_order <- rownames(J)
  
  expand.grid(recipient = sp_order, actor = sp_order,
              stringsAsFactors = FALSE) %>%
    as_tibble() %>%
    filter(recipient != actor) %>%
    mutate(
      interaction_strength = map2_dbl(recipient, actor, ~ J[.x, .y]),
      direction = case_when(
        interaction_strength > 0 ~ "positive",
        interaction_strength < 0 ~ "negative",
        TRUE                     ~ "zero"
      )
    ) %>%
    filter(interaction_strength != 0)
}

```





```{r}
# =============================================================================
# CONSTANTS
# =============================================================================

basal_taxa <- c("amorphous_detritus", "animal", "cyanobacteria",
                "diatom", "filamentous", "green_algae", "plant_material")

# Season assignment, based on GCA timing: first = Jan-Apr, second = May-Aug, third = Sep-Dec
community <- community %>%
  mutate(season = case_when(
    month(date_id) %in% 1:4  ~ "first",
    month(date_id) %in% 5:8  ~ "second",
    month(date_id) %in% 9:12 ~ "third"
  ))

seasons <- c("first", "second", "third")
sites   <- names(diet)

# =============================================================================
# MASTER LOOP: all sites x annual + seasonal
# =============================================================================

results <- map(sites, function(s) {
  
  message("\n=== Site: ", s, " ===")
  
  # --- ANNUAL ---
  message("  Annual...")
  diet_ann     <- diet[[s]][["annual"]]
  comm_ann     <- prep_community(community, s, season_name = NULL, diet_ann)
  annual_out   <- if (!is.null(comm_ann)) process_one_scale(comm_ann, diet_ann) else NULL
  
  # --- SEASONAL ---
  seasonal_out <- map(seasons, function(ss) {
    message("  Season: ", ss)
    diet_seas  <- diet[[s]][["season_means"]][[ss]]
    if (is.null(diet_seas)) return(NULL)
    comm_seas  <- prep_community(community, s, ss, diet_seas)
    if (is.null(comm_seas)) return(NULL)
    process_one_scale(comm_seas, diet_seas)
  }) %>% set_names(seasons)
  
  list(annual = annual_out, seasonal = seasonal_out)
  
}) %>% set_names(sites)

# =============================================================================
# FLATTEN TO ANALYSIS-READY TIBBLES
# =============================================================================

# Helper to safely extract with labels
extract_df <- function(results, element) {
  map_dfr(sites, function(s) {
    
    ann <- results[[s]]$annual[[element]]
    if (!is.null(ann)) ann <- ann %>%
        mutate(site = s, scale = "annual", season = NA_character_)
    
    seas <- map_dfr(seasons, function(ss) {
      x <- results[[s]]$seasonal[[ss]][[element]]
      if (!is.null(x)) x %>%
          mutate(site = s, scale = "seasonal", season = ss)
    })
    
    bind_rows(ann, seas)
  })
}

stability_all     <- extract_df(results, "stability_summary")
interactions_all  <- extract_df(results, "interaction_strengths")

```


# Stability visualizations
```{r}
head(stability_all)

hist(stability_all$resilience)
hist(stability_all$reactivity)




```

Visualizing the interaction strengths from the off-diagonal Jacobian output - the per-capita effect of each species on another. Values are incredibly large because of the 0gm-2 basal biomass assumption.
```{r}

# Base R skewness function
skewness_base <- function(x) {
  x <- x[is.finite(x)]
  n <- length(x)
  m <- mean(x)
  s <- sd(x)
  (sum((x - m)^3) / n) / s^3
}

library(ggplot2)
library(dplyr)
library(patchwork)

site_order  <- c("st14", "oh2", "st7", "st9", "st6", "hver")
site_colors <- c(st14="#2C7BB6", oh2="#5AAE61", st7="#A6D96A",
                 st9="#FDAE61", st6="#F46D43", hver="#D73027")

int_sum <- interactions_all %>%
  group_by(site, scale, season) %>%
  summarise(
    mean_abs = mean(abs(interaction_strength)),
    var_abs  = var(abs(interaction_strength)),
    skew_abs = skewness_base(abs(interaction_strength)),
    cv_abs   = sd(abs(interaction_strength)) / mean(abs(interaction_strength)),
    .groups  = "drop"
  ) %>%
  mutate(
    site = factor(site, levels = site_order),
    time_label = case_when(
      scale == "annual"   ~ "Annual",
      season == "first"   ~ "Jan–Apr",
      season == "second"  ~ "May–Aug",
      season == "third"   ~ "Sep–Dec"
    ),
    time_label = factor(time_label, 
                        levels = c("Annual","Jan–Apr","May–Aug","Sep–Dec"))
  )

# Label which direction each J_ij represents
interactions_all <- interactions_all %>%
  mutate(j_type = case_when(
    interaction_strength > 0 ~ "resource_effect_on_consumer",  # artifact tail
    interaction_strength < 0 ~ "consumer_effect_on_resource"   # real signal
  ))

# Analyze only the ecologically meaningful half
interactions_real <- interactions_all %>%
  filter(j_type == "consumer_effect_on_resource")

# --- Plot 1: Skewness across gradient (most interpretable) ---
p_skew <- ggplot(int_sum %>% filter(scale == "seasonal"),
                 aes(x = time_label, y = skew_abs, 
                     color = site, group = site)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 3) +
  scale_color_manual(values = site_colors) +
  labs(x = NULL, y = "Skewness of |J_ij|",
       title = "Interaction strength skewness",
       subtitle = "Higher = more dominated by rare strong interactions") +
  theme_classic(base_size = 13)

# --- Plot 2: CV as normalized dispersion (avoids scale artifact) ---
p_cv <- ggplot(int_sum %>% filter(scale == "seasonal"),
               aes(x = time_label, y = cv_abs,
                   color = site, group = site)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 3) +
  scale_color_manual(values = site_colors) +
  labs(x = NULL, y = "CV of |J_ij|",
       title = "Coefficient of variation",
       subtitle = "Scale-free dispersion of interaction strengths") +
  theme_classic(base_size = 13)

# --- Plot 3: Ridge plot of raw |J_ij| distribution per site (annual) ---
library(ggridges)

p_ridge <- interactions_all %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = rev(site_order)),
         log_abs_J = log10(abs(interaction_strength) + 1)) %>%
  ggplot(aes(x = log_abs_J, y = site, fill = site)) +
  geom_density_ridges(alpha = 0.8, scale = 1.2) +
  scale_fill_manual(values = site_colors) +
  labs(x = expression(log[10]("|J"[ij]*"|")),
       y = NULL,
       title = "Distribution of interaction strengths (annual)",
       subtitle = "Cold (top) → warm (bottom)") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

cr_annual <- interactions_real %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = rev(site_order)),
         log_abs_J = log10(abs(interaction_strength) + 1)) %>%
  ggplot(aes(x = log_abs_J, y = site, fill = site)) +
  geom_density_ridges(alpha = 0.8, scale = 1.2) +
  scale_fill_manual(values = site_colors) +
  labs(x = expression(log[10]("|J"[ij]*"|")),
       y = NULL,
       title = "Consumer-resource interaction strengths (annual) - off-diagonal Jacobian") +
  theme_classic(base_size = 13)

cr_annual_facetResource <-interactions_real %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = rev(site_order)),
         log_abs_J = log10(abs(interaction_strength) + 1)) %>%
  ggplot(aes(x = log_abs_J, y = site, fill = site)) +
  geom_density_ridges(alpha = 0.8, scale = 1.2) +
  scale_fill_manual(values = site_colors) +
  labs(x = expression(log[10]("|J"[ij]*"|")),
       y = NULL,
       title = "Consumer-resource interaction strengths (annual) - off-diagonal Jacobian, by resource") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none") +
  facet_wrap(~recipient, ncol = 3)


# Save tiff
ggsave(
  filename    = "figures/CRinteraction.tiff",
  plot        = cr_annual,
  width       = 8,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"  
)

# Save tiff
ggsave(
  filename    = "figures/CRinteraction_facet.tiff",
  plot        = cr_annual_facetResource,
  width       = 8,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"  
)

ggsave(
  filename = "figures/CRinteraction_facet.pdf",
  plot     = cr_annual_facetResource,
  width    = 6,
  height   = 4.5,
  units    = "in"
)

ggsave(
  filename = "figures/CRinteraction.pdf",
  plot     = cr_annual,
  width    = 6,
  height   = 4.5,
  units    = "in"
)



# --- Plot 4: Mean vs skewness stability space ---
p_space <- ggplot(int_sum %>% filter(scale == "annual"),
                  aes(x = log10(mean_abs), y = skew_abs, 
                      color = site, label = site)) +
  geom_point(size = 4) +
  ggrepel::geom_text_repel(size = 4) +
  scale_color_manual(values = site_colors) +
  labs(x = expression(log[10]("mean |J"[ij]*")")),
       y = "Skewness of |J_ij|",
       title = "Interaction strength space") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

(p_skew | p_cv) / (p_ridge | p_space)


# Pull annual Jacobian for st6 and find the strong interactions
J_st6 <- results[["st6"]]$annual$jacobians[[1]]  # first bootstrap as example

# Find the outlier pairs driving the tail
as.data.frame(J_st6) %>%
  rownames_to_column("recipient") %>%
  pivot_longer(-recipient, names_to = "actor", values_to = "J") %>%
  filter(recipient != actor, J != 0) %>%
  arrange(desc(abs(J))) %>%
  head(20)
```




```{r}
# =============================================================================
# EXTRACT FLUX MATRICES → long tidy tibble
# columns: site, scale, season, boot_id, from (prey), to (consumer), flux_J
# =============================================================================

extract_fluxes <- function(results) {
  map_dfr(sites, function(s) {
    
    # Annual
    ann_fl <- results[[s]]$annual$flux_list
    ann_df <- if (!is.null(ann_fl)) {
      map_dfr(names(ann_fl), function(b) {
        mat <- ann_fl[[b]]
        as.data.frame(as.table(mat)) %>%          # melt matrix to long
          setNames(c("prey", "consumer", "flux_J")) %>%
          filter(flux_J > 0) %>%                  # drop true zeros
          mutate(boot_id = b, site = s, 
                 scale = "annual", season = NA_character_)
      })
    }
    
    # Seasonal
    seas_df <- map_dfr(seasons, function(ss) {
      fl <- results[[s]]$seasonal[[ss]]$flux_list
      if (is.null(fl)) return(NULL)
      map_dfr(names(fl), function(b) {
        mat <- fl[[b]]
        as.data.frame(as.table(mat)) %>%
          setNames(c("from", "to", "flux_J")) %>%
          filter(flux_J > 0) %>%
          mutate(boot_id = b, site = s, 
                 scale = "seasonal", season = ss)
      })
    })
    
    bind_rows(ann_df, seas_df)
  })
}

flux_all    <- extract_fluxes(results)

# =============================================================================
# EXTRACT JACOBIANS → long tidy tibble
# columns: site, scale, season, boot_id, actor, recipient, j_value
# =============================================================================

extract_jacobians <- function(results) {
  map_dfr(sites, function(s) {
    
    # Annual
    ann_jac <- results[[s]]$annual$jacobians
    ann_df  <- if (!is.null(ann_jac)) {
      map_dfr(names(ann_jac), function(b) {
        J <- ann_jac[[b]]
        as.data.frame(as.table(J)) %>%
          setNames(c("recipient", "actor", "j_value")) %>%
          mutate(boot_id = b, site = s, 
                 scale = "annual", season = NA_character_)
      })
    }
    
    # Seasonal
    seas_df <- map_dfr(seasons, function(ss) {
      jac <- results[[s]]$seasonal[[ss]]$jacobians
      if (is.null(jac)) return(NULL)
      map_dfr(names(jac), function(b) {
        J <- jac[[b]]
        as.data.frame(as.table(J)) %>%
          setNames(c("recipient", "actor", "j_value")) %>%
          mutate(boot_id = b, site = s, 
                 scale = "seasonal", season = ss)
      })
    })
    
    bind_rows(ann_df, seas_df)
  })
}


jacobian_all <- extract_jacobians(results)


site_colors <- c(
  st14 = "#2C7BB6",  # deep blue (coldest)
  oh2  = "#5AAE61",  # blue-green
  st7  = "#A6D96A",  # green
  st9  = "#FDAE61",  # orange
  st6  = "#F46D43",  # orange-red
  hver = "#D73027"   # red (warmest)
)

site_order <- c("st14", "oh2", "st7", "st9", "st6", "hver")

# --- Flux: distribution of log10 fluxes per site (annual) ---
flux_all %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = log10(flux_J), y = site, fill = site)) +
  geom_density_ridges(alpha = 0.8) +
  scale_fill_manual(values = site_colors) +
  labs(
    x = expression(log[10](Energy~flux~(J~m^{-2}~day^{-1}))),
    y = NULL,
    title = "Distribution of energy fluxes by site (annual)"
  ) +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")


st14 <- flux_all %>%
  filter(site == "st14")

# Seasonal fluxes
flux_all %>%
  filter(scale == "seasonal") %>%
  mutate(
    site = factor(site, levels = site_order),
    season_label = case_when(
      season == "first"  ~ "Jan–Apr",
      season == "second" ~ "May–Aug",
      season == "third"  ~ "Sep–Dec"
    ),
    season_label = factor(season_label, levels = c("Jan–Apr", "May–Aug", "Sep–Dec"))
  ) %>%
  ggplot(aes(x = season_label, y = log(flux_J), fill = site)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_manual(values = site_colors) +
  labs(
    y     = expression(ln(Energy~flux~(J~m^{-2}~day^{-1}))),
    x     = "Season",
    title = "Seasonal energy flux distributions across the thermal gradient"
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 12),
    axis.text.y      = element_text(size = 10)
  )



# Seasonal visualizations 
flux_seasonal_summary <- flux_all %>%
  filter(scale == "seasonal") %>%
  mutate(
    site = factor(site, levels = site_order),
    season_label = case_when(
      season == "first"  ~ "Jan–Apr",
      season == "second" ~ "May–Aug",
      season == "third"  ~ "Sep–Dec"
    ),
    season_label = factor(season_label, levels = c("Jan–Apr", "May–Aug", "Sep–Dec"))
  ) %>%
  # Step 1: total flux per bootstrap (collapse across links)
  group_by(site, season_label, boot_id) %>%
  summarise(total_flux = sum(flux_J), .groups = "drop") %>%
  # Step 2: summarise across bootstraps
  group_by(site, season_label) %>%
  summarise(
    mean_flux   = mean(total_flux),
    median_flux = median(total_flux),
    lo_95       = quantile(total_flux, 0.025),
    hi_95       = quantile(total_flux, 0.975),
    lo_50       = quantile(total_flux, 0.25),    # inner 50% if you want to show IQR too
    hi_50       = quantile(total_flux, 0.75),
    .groups = "drop"
  )


# Annual summary (same structure as seasonal)
flux_annual_summary <- flux_all %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  group_by(site, boot_id) %>%
  summarise(total_flux = sum(flux_J), .groups = "drop") %>%
  group_by(site) %>%
  summarise(
    mean_flux   = mean(total_flux),
    median_flux = median(total_flux),
    lo_95       = quantile(total_flux, 0.025),
    hi_95       = quantile(total_flux, 0.975),
    .groups     = "drop"
  ) %>%
  mutate(season_label = "Annual")

# Derive annual from seasonal by averaging across seasons per bootstrap
flux_annual_from_seasonal <- flux_all %>%
  filter(scale == "seasonal") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  # Step 1: total flux per site x season x boot
  group_by(site, season, boot_id) %>%
  summarise(total_flux = sum(flux_J), .groups = "drop") %>%
  # Step 2: average across seasons within each boot
  group_by(site, boot_id) %>%
  summarise(annual_flux = mean(total_flux), .groups = "drop") %>%
  # Step 3: summarise across boots
  group_by(site) %>%
  summarise(
    mean_flux   = mean(annual_flux),
    median_flux = median(annual_flux),
    lo_95       = quantile(annual_flux, 0.025),
    hi_95       = quantile(annual_flux, 0.975),
    .groups     = "drop"
  ) %>%
  mutate(season_label = "Annual")

# Rebuild combined plot with corrected annual
flux_combined <- flux_seasonal_summary %>%
  bind_rows(flux_annual_from_seasonal) %>%
  mutate(
    season_label = factor(season_label,
                          levels = c("Jan–Apr", "May–Aug", "Sep–Dec", "Annual")),
    site = factor(site, levels = site_order)
  )

p_mean <- ggplot(flux_combined,
                 aes(x = season_label, y = mean_flux,
                     fill = site, group = site)) +  # fill instead of colour
  geom_vline(xintercept = 3.5, linetype = "dashed",
             colour = "grey70", linewidth = 0.5) +
  geom_pointrange(
    aes(ymin = lo_95, ymax = hi_95, colour = site),
    position = position_dodge(width = 0.3),
    size     = 0.6,
    colour   = "black",    # outline colour
    linewidth = 0.5
  ) +
  geom_point(
    aes(fill = site),
    position = position_dodge(width = 0.3),
    shape    = 21,         # filled circle with outline
    size     = 3,
    colour   = "black",    # outline colour
    stroke   = 0.6         # outline thickness
  ) +
  geom_smooth(method = "lm") +
  scale_fill_manual(values = site_colors) +
  scale_colour_manual(values = site_colors, guide = "none") +  # hide duplicate legend
  scale_y_log10() +
  labs(
    x    = NULL,
    y    = expression(Total~energy~flux~(J~m^{-2}~d^{-1})),
    fill = "Site"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.text.x = element_text(
      face = c("plain", "plain", "plain", "bold")
    )
  )

p_mean


# Save tiff
ggsave(
  filename    = "figures/seasonal_energy_flux.tiff",
  plot        = p_mean,
  width       = 8,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"  
)

# Save pdf
ggsave(
  filename = "figures/seasonal_energy_flux.pdf",
  plot     = p_mean,
  width    = 6,
  height   = 4.5,
  units    = "in"
)


flux_change <- flux_combined %>%
  filter(season_label != "Annual") %>%          # seasonal only
  group_by(site) %>%
  mutate(
    # Use Jan-Apr as baseline
    baseline_median = median_flux[season_label == "Jan–Apr"],
    baseline_mean   = mean_flux[season_label == "Jan–Apr"],
    
    # % change from Jan-Apr baseline
    pct_change_median = ((median_flux - baseline_median) / baseline_median) * 100,
    pct_change_mean   = ((mean_flux   - baseline_mean)   / baseline_mean)   * 100
  ) %>%
  ungroup()

# Quick summary of peak season (May-Aug) % change by site
flux_change %>%
  filter(season_label == "May–Aug") %>%
  select(site, baseline_median, median_flux, pct_change_median) %>%
  arrange(desc(pct_change_median))




# What drives the energy flux variation across the thermal gradient?

# Correlate boot-level total flux with boot-level total biomass per site x season
flux_all %>%
  filter(scale == "seasonal") %>%
  group_by(site, season, boot_id) %>%
  summarise(total_flux = sum(flux_J), .groups = "drop") %>%
  left_join(
    community %>%
      filter(!taxon_id %in% basal_taxa) %>%
      group_by(site, season, boot_id) %>%
      summarise(total_biomass = sum(bio_g_m),
                mean_metrate  = mean(metrate_J_day_gram, na.rm = TRUE),
                mean_bodysize = mean(ind_g, na.rm = TRUE),
                .groups = "drop"),
    by = c("site", "season", "boot_id")
  ) %>%
  ggplot(aes(x = log(mean_metrate), y = log(total_flux), colour = site)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~ season) +
  theme_classic()


# Site-level annual means — flux vs temperature
flux_annual_from_seasonal %>%
  left_join(
    community %>%
      group_by(site) %>%
      summarise(
        mean_temp_C   = mean(celcius, na.rm = TRUE),
        mean_bodysize = mean(ind_g[!taxon_id %in% basal_taxa], na.rm = TRUE),
        total_biomass = mean(bio_g_m[!taxon_id %in% basal_taxa], na.rm = TRUE)
      ),
    by = "site"
  ) %>%
  ggplot(aes(x = mean_temp_C, y = mean_flux, colour = site, label = site)) +
  geom_pointrange(aes(ymin = lo_95, ymax = hi_95)) +
  geom_smooth(aes(group = 1), method = "lm", colour = "black", se = TRUE) +
  ggrepel::geom_text_repel(size = 3) +
  scale_colour_manual(values = site_colors) +
  scale_y_log10() +
  theme_classic() +
  labs(x = "Mean annual temperature (°C)",
       y = expression(Total~energy~flux~(J~m^{-2}~d^{-1})))


# Community mean body size vs temperature
community %>%
  filter(!taxon_id %in% basal_taxa, ind_g > 0) %>%
  group_by(site, season) %>%
  summarise(
    mean_bodysize = mean(ind_g, na.rm = TRUE),
    mean_temp     = mean(celcius, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = mean_temp, y = log(mean_bodysize),
             colour = site, label = site)) +
  geom_point(size = 3) +
  geom_smooth(aes(group = 1), method = "lm", colour = "black") +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~ season) +
  theme_classic() +
  labs(x = "Mean temperature (°C)",
       y = "ln(Mean body size, g/ind)")

```

```{r}
library(lme4)
library(MuMIn)

model_df <- flux_all %>%
  filter(scale == "seasonal") %>%
  group_by(site, season, boot_id) %>%
  summarise(total_flux = sum(flux_J), .groups = "drop") %>%
  left_join(
    community %>%
      filter(
        !taxon_id %in% basal_taxa,
        bio_g_m > 0,          # exclude placeholder zeros
        ind_g > 1e-6,         # exclude nonsensical body sizes
        n_ind_m > 0           # exclude zero abundance
      ) %>%
      group_by(site, season, boot_id) %>%
      summarise(
        total_biomass = sum(bio_g_m, na.rm = TRUE),
        # biomass-weighted mean metabolic rate
        mean_metrate  = weighted.mean(metrate_J_day_gram, 
                                      w = bio_g_m, na.rm = TRUE),
        # biomass-weighted mean body size
        mean_bodysize = weighted.mean(ind_g, 
                                      w = bio_g_m, na.rm = TRUE),
        mean_temp     = mean(celcius, na.rm = TRUE),
        .groups = "drop"),
    by = c("site", "season", "boot_id")
  ) %>%
  mutate(across(c(total_flux, total_biomass,
                  mean_metrate, mean_bodysize), log))

head(model_df)

# Quick sanity check — back-transform to original units
model_df %>%
  group_by(site) %>%
  summarise(across(c(total_flux, total_biomass, mean_metrate, mean_bodysize),
                   ~ mean(exp(.x)))) %>%
  arrange(site)

bs<- model_df %>%
  mutate(site = factor(site, levels = site_order)) %>%
  group_by(site) %>%
  summarise(
    mean_bs   = mean(exp(mean_bodysize)) * 1000,
    lo_95     = quantile(exp(mean_bodysize), 0.025) * 1000,
    hi_95     = quantile(exp(mean_bodysize), 0.975) * 1000,
    mean_temp = mean(mean_temp)
  ) %>%
  mutate(site = factor(site, levels = site_order)) %>%  # re-apply after summarise
  ggplot(aes(x = mean_temp, y = mean_bs, colour = site, label = site)) +
  geom_pointrange(
    aes(ymin = lo_95, ymax = hi_95, colour = site),
    position = position_dodge(width = 0.3),
    size     = 0.6,
    colour   = "black",    # outline colour
    linewidth = 0.5
  ) +
  geom_point(
    aes(fill = site),
    position = position_dodge(width = 0.3),
    shape    = 21,         # filled circle with outline
    size     = 3.7,
    colour   = "black",    # outline colour
    stroke   = 0.6         # outline thickness
  ) +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 30, by = 5)) +
  scale_fill_manual(values = site_colors, name = "Site") +
  theme_classic(base_size = 13) +
  labs(
    x = "Mean annual temperature (°C)",
    y = expression(Community~body~size~(mg~ind^{-1}))
  )

bs
# Save tiff
ggsave(
  filename    = "figures/bodysize_temp.tiff",
  plot        = bs,
  width       = 7,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"
)

# Save pdf
ggsave(
  filename    = "figures/bodysize_temp.pdf",
  plot        = bs,
  width    = 7,
  height   = 4.5,
  units    = "in"
)

```

```{r}
# =============================================================================
# FIGURE 1: Taxon-level flux contributions — stacked bars
# Which consumers dominate total flux at each site, and does composition shift?
# Built from annual flux_list — colSums per taxon per boot, then summarised
# =============================================================================

taxon_flux_annual <- map_dfr(sites, function(s) {
  fl <- results[[s]]$annual$flux_list
  if (is.null(fl)) return(NULL)
  
  map_dfr(names(fl), function(b) {
    mat <- fl[[b]]
    colSums(mat) %>%
      enframe(name = "taxon_id", value = "flux_in") %>%
      mutate(site = s, boot_id = b)
  })
})

# Summarise: mean flux per taxon per site across boots
taxon_flux_summary <- taxon_flux_annual %>%
  filter(!taxon_id %in% basal_taxa, flux_in > 0) %>%
  group_by(site, taxon_id) %>%
  summarise(
    mean_flux  = mean(flux_in,   na.rm = TRUE),
    total_flux = sum(flux_in,    na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(site) %>%
  mutate(
    site_total   = sum(mean_flux),
    pct_of_total = mean_flux / site_total * 100
  ) %>%
  ungroup() %>%
  mutate(site = factor(site, levels = site_order))

# Keep only taxa that contribute > 10% at any site; lump rest as "Other"
dominant_taxa <- taxon_flux_summary %>%
  group_by(taxon_id) %>%
  summarise(max_pct = max(pct_of_total)) %>%
  filter(max_pct > 5) %>%
  pull(taxon_id)

# --- Build plot dataframe with absolute flux per taxon ---
taxon_flux_abs <- taxon_flux_summary %>%
  mutate(taxon_id = if_else(taxon_id %in% dominant_taxa, taxon_id, "Other")) %>%
  group_by(site, taxon_id) %>%
  summarise(mean_flux = sum(mean_flux), .groups = "drop") %>%
  mutate(site = factor(site, levels = site_order))

# --- Count number of consumer taxa per site (before lumping) ---
n_taxa_site <- taxon_flux_summary %>%
  filter(taxon_id != "Other") %>%
  group_by(site) %>%
  summarise(n_taxa = n_distinct(taxon_id), .groups = "drop") %>%
  mutate(site = factor(site, levels = site_order))

# --- Total flux per site for label positioning ---
site_totals <- taxon_flux_abs %>%
  group_by(site) %>%
  summarise(total = sum(mean_flux), .groups = "drop") %>%
  left_join(n_taxa_site, by = "site")

# --- Reorder taxon fill so "Other" always at bottom ---
taxon_order <- c("Other", sort(dominant_taxa))
taxon_flux_abs <- taxon_flux_abs %>%
  mutate(taxon_id = factor(taxon_id, levels = taxon_order))

# --- Combined figure ---
p_combined <- ggplot(taxon_flux_abs,
                     aes(x = site, y = mean_flux, fill = taxon_id)) +
  geom_col(colour = "black", linewidth = 0.3, position = "stack") +
  # Annotate n taxa above each bar
  geom_text(
    data     = site_totals,
    aes(x = site, y = total, 
        label = paste0("n = ", n_taxa)),
    inherit.aes = FALSE,
    vjust   = -0.5,
    size    = 3.5,
    colour  = "black"
  ) +
  scale_fill_manual(values = tax_cols, name = "Consumer taxon") +
  scale_y_continuous(limits = c(0,80), expand = c(0,0),
    labels = scales::label_comma()
  ) +
  labs(
    x = NULL,
    y = expression(Mean~total~energy~flux~(J~m^{-2}~d^{-1})),
    title    = "Community energy flux across the thermal gradient",
  ) +
  theme_classic(base_size = 13) +
  theme(
    legend.text      = element_text(size = 9, face = "italic"),
    legend.key.size  = unit(0.4, "cm"),
    legend.title     = element_text(size = 10),
    legend.position  = "right",
    plot.subtitle    = element_text(size = 10, colour = "grey40"),
    axis.text.x      = element_text(size = 11),
    plot.margin      = margin(t = 5, r = 5, b = 25, l = 5)  # room for arrow
  )

p_combined

# --- Save ---
ggsave(
  filename    = "figures/flux_combined.tiff",
  plot        = p_combined,
  width       = 9,
  height      = 5.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"
)

ggsave(
  filename = "figures/flux_combined.pdf",
  plot     = p_combined,
  width    = 9,
  height   = 5.5,
  units    = "in"
)


# =============================================================================
# FIGURE 2: Mass-specific flux across the thermal gradient
# Total flux / total community biomass — separates biomass from metabolic signal
# =============================================================================

# Total biomass per site x boot (annual, consumers only)
biomass_annual <- community %>%
  filter(!taxon_id %in% basal_taxa, bio_g_m > 0) %>%
  group_by(site, boot_id) %>%
  summarise(total_bio = sum(bio_g_m, na.rm = TRUE), .groups = "drop")

# Total flux per site x boot (annual)
total_flux_annual <- taxon_flux_annual %>%
  filter(!taxon_id %in% basal_taxa) %>%
  group_by(site, boot_id) %>%
  summarise(total_flux = sum(flux_in, na.rm = TRUE), .groups = "drop")

# Mass-specific flux = total flux / total biomass (J g-1 day-1)
ms_flux <- total_flux_annual %>%
  left_join(biomass_annual, by = c("site", "boot_id")) %>%
  mutate(ms_flux = total_flux / total_bio) %>%
  group_by(site) %>%
  summarise(
    mean_ms   = mean(ms_flux,              na.rm = TRUE),
    lo_95     = quantile(ms_flux, 0.025,   na.rm = TRUE),
    hi_95     = quantile(ms_flux, 0.975,   na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  mutate(site = factor(site, levels = site_order))

# Join annual temperature for x-axis
site_temps <- community %>%
  group_by(site) %>%
  summarise(mean_temp_c = mean(celcius, na.rm = TRUE), .groups = "drop")

ms_flux <- ms_flux %>%
  left_join(site_temps, by = "site")

p_ms_flux <- ggplot(ms_flux,
                    aes(x = mean_temp_c, y = mean_ms, colour = site)) +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 0.8, se = TRUE) +
  geom_pointrange(aes(ymin = lo_95, ymax = hi_95), size = 0.8) +
  scale_colour_manual(values = site_colors) +
  scale_y_log10() +
  labs(
    x      = "Mean annual temperature (°C)",
    y      = expression(Mass-specific~flux~(J~g^{-1}~d^{-1})),
    colour = "Site",
    title  = "Mass-specific energy flux across the thermal gradient",
    subtitle = "Separates metabolic rate effect from standing biomass effect"
  ) +
  theme_classic(base_size = 13)

# Print all three
p_stacked
p_ms_flux

```



```{r}
focal_taxon <- "Simulium vittatum"  # swap to whatever coverage check reveals

# --- Pull biomass and metabolic rate from community ---
taxon_community <- community %>%
  filter(taxon_id == focal_taxon, !taxon_id %in% basal_taxa) %>%
  mutate(site = factor(site, levels = site_order))

# --- Pull flux received by this taxon from flux_all ---
taxon_flux <- flux_all %>%
  filter(consumer == focal_taxon, scale == "annual") %>%
  group_by(site, boot_id) %>%
  summarise(total_flux_in = sum(flux_J), .groups = "drop") %>%
  mutate(site = factor(site, levels = site_order))

# --- Summarise across bootstraps ---
bio_summary <- taxon_community %>%
  group_by(site) %>%
  filter(bio_g_m > 0) %>%
  summarise(
    mean_val = mean(bio_g_m, na.rm = TRUE),
    lo_95    = quantile(bio_g_m, 0.025, na.rm = TRUE),
    hi_95    = quantile(bio_g_m, 0.975, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(metric = "Biomass (g m-2)")

met_summary <- taxon_community %>%
  group_by(site) %>%
  filter(metrate_J_day_gram > 0) %>%
  summarise(
    mean_val = mean(metrate_J_day_gram, na.rm = TRUE),
    lo_95    = quantile(metrate_J_day_gram, 0.025, na.rm = TRUE),
    hi_95    = quantile(metrate_J_day_gram, 0.975, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(metric = "Metabolic rate (J g-1 d-1)")

bodysize_summary <- taxon_community %>%
  group_by(site) %>%
  filter(ind_g > 0) %>%
  summarise(
    mean_val = mean(ind_g, na.rm = TRUE),
    lo_95    = quantile(ind_g, 0.025, na.rm = TRUE),
    hi_95    = quantile(ind_g, 0.975, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(metric = "Body size (g ind-1)")

flux_summary <- taxon_flux %>%
  group_by(site) %>%
  filter(total_flux_in > 0) %>%
  summarise(
    mean_val = mean(total_flux_in, na.rm = TRUE),
    lo_95    = quantile(total_flux_in, 0.025, na.rm = TRUE),
    hi_95    = quantile(total_flux_in, 0.975, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(metric = "Energy flux in (J m-2 d-1)")

# --- Combine and plot ---
combined <- bind_rows(bio_summary, met_summary, bodysize_summary, flux_summary) %>%
  mutate(
    metric = factor(metric, levels = c(
      "Biomass (g m-2)",
      "Metabolic rate (J g-1 d-1)",
      "Body size (g ind-1)",
      "Energy flux in (J m-2 d-1)"
    )),
    site = factor(site, levels = site_order)
  )


ggplot(combined, aes(x=site,y=mean_val, fill = site))+
  geom_pointrange(aes(ymin = lo_95, ymax = hi_95, color = site),
                  shape = 21,
                  stroke = 0.6,
                  size = 0.7) +
  scale_fill_manual(values   = site_colors, guide = "none") +
  scale_colour_manual(values = site_colors, guide = "none") +
  scale_y_log10() +
  facet_wrap(~metric) +
  labs(
    x        = NULL,
    y        = "Site") +
  theme_classic(base_size = 13) +
  theme(
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 11),
    legend.position  = "none",
    axis.text.x      = element_text(angle = 30, hjust = 1, size = 10),
    panel.spacing    = unit(1, "lines")
  )


# --- Shared theme ---
taxon_theme <- theme_classic(base_size = 13) +
  theme(
    legend.position = "none",
    axis.text.x     = element_text(angle = 30, hjust = 1, size = 10),
    panel.spacing   = unit(1, "lines")
  )

# --- Helper to build each panel ---
make_panel <- function(data, y_label) {
  ggplot(data,
         aes(x = site, y = mean_val, fill = site)) +
    geom_pointrange(
      aes(ymin = lo_95, ymax = hi_95, colour = site),
      shape  = 21,
      stroke = 0.6,
      size   = 0.7
    ) +
    scale_fill_manual(values   = site_colors, guide = "none") +
    scale_colour_manual(values = site_colors, guide = "none") +
    scale_y_log10() +
    scale_x_discrete(limits = site_order) +
    labs(x = NULL,
         y = y_label) +
    taxon_theme
}

# --- Build each panel with correct y label ---
p_bio <- make_panel(
  bio_summary %>% mutate(site = factor(site, levels = site_order)),
  expression(Biomass~(g~m^{-2}))
)

p_met <- make_panel(
  met_summary %>% mutate(site = factor(site, levels = site_order)),
  expression(Metabolic~rate~(J~g^{-1}~d^{-1}))
)

p_size <- make_panel(
  bodysize_summary %>% mutate(site = factor(site, levels = site_order)),
  expression(Body~size~(g~ind^{-1}))
)

p_flux <- make_panel(
  flux_summary %>% mutate(site = factor(site, levels = site_order)),
  expression(Energy~flux~(J~m^{-2}~d^{-1}))
)

# --- Combine ---
p_taxon <- (p_bio | p_met) / (p_size | p_flux) +
  plot_annotation(
    title    = paste("Metabolic regime of", focal_taxon,
                     "across the thermal gradient"),
    theme    = theme(
      plot.title    = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 10, colour = "grey40")
    )
  )

p_taxon

# --- Save ---
ggsave(
  filename    = paste0("figures/focal_taxon_", focal_taxon, ".tiff"),
  plot        = p_taxon,
  width       = 8,
  height      = 7,
  units       = "in",
  dpi         = 300,
  compression = "lzw"
)

ggsave(
  filename = paste0("figures/focal_taxon_", focal_taxon, ".pdf"),
  plot     = p_taxon,
  width    = 8,
  height   = 7,
  units    = "in"
)
# --- Save ---
ggsave(
  filename    = paste0("figures/focal_taxon_", focal_taxon, ".tiff"),
  plot        = p_taxon,
  width       = 6,
  height      = 7,
  units       = "in",
  dpi         = 300,
  compression = "lzw"
)
```

```{r}
# =============================================================================
# TURNOVER–STABILITY ANALYSIS
# Does community-level energy turnover rate predict resilience and reactivity?
# =============================================================================

# -----------------------------------------------------------------------------
# STEP 1: Extract total flux INTO each consumer, per boot x site x scale
# Column sums of flux matrix = energy flowing from all resources to consumer i
# -----------------------------------------------------------------------------

turnover_boot <- map_dfr(sites, function(s) {
  
  # Annual
  fl_ann <- results[[s]]$annual$flux_list
  ann_out <- if (!is.null(fl_ann)) {
    map_dfr(names(fl_ann), function(b) {
      colSums(fl_ann[[b]]) %>%
        enframe(name = "taxon_id", value = "flux_in") %>%
        mutate(site = s, scale = "annual", season = "annual", boot_id = b)
    })
  }
  
  # Seasonal
  seas_out <- map_dfr(seasons, function(ss) {
    fl <- results[[s]]$seasonal[[ss]]$flux_list
    if (is.null(fl)) return(NULL)
    map_dfr(names(fl), function(b) {
      colSums(fl[[b]]) %>%
        enframe(name = "taxon_id", value = "flux_in") %>%
        mutate(site = s, scale = "seasonal", season = ss, boot_id = b)
    })
  })
  
  bind_rows(ann_out, seas_out)
})

# -----------------------------------------------------------------------------
# STEP 2: Pull boot-level consumer biomass from community df
# Replicates the aggregation logic in prep_community():
#   annual  = mean bio_g_m across all dates per boot x taxon
#   seasonal = mean bio_g_m within season per boot x taxon
# Use "annual" as sentinel (avoids NA-join issues with season column)
# -----------------------------------------------------------------------------

biomass_boot <- bind_rows(
  # Annual: mean across all dates
  community %>%
    group_by(site, boot_id, taxon_id) %>%
    summarise(bio_g_m = mean(bio_g_m, na.rm = TRUE), .groups = "drop") %>%
    mutate(scale = "annual", season = "annual"),
  
  # Seasonal: mean within each season
  community %>%
    group_by(site, boot_id, taxon_id, season) %>%
    summarise(bio_g_m = mean(bio_g_m, na.rm = TRUE), .groups = "drop") %>%
    mutate(scale = "seasonal")
)

# -----------------------------------------------------------------------------
# STEP 3: Join flux with biomass, compute per-consumer turnover rate
# Turnover (day-1) = flux into consumer / consumer biomass
# Exclude basal taxa — placeholder biomasses (~0) make their rates meaningless
# -----------------------------------------------------------------------------

turnover_rates <- turnover_boot %>%
  left_join(biomass_boot,
            by = c("site", "boot_id", "taxon_id", "scale", "season")) %>%
  filter(
    !taxon_id %in% basal_taxa,   # basal biomasses are placeholders
    bio_g_m    > 0,              # guard against residual zeros
    flux_in    > 0               # only fed consumers
  ) %>%
  mutate(turnover_rate = flux_in / bio_g_m)

# -----------------------------------------------------------------------------
# STEP 4: Community-level summary per boot
# Biomass-weighted geometric mean turnover — weights large populations more
# heavily, geometric mean appropriate because rates are log-normally distributed
# -----------------------------------------------------------------------------

turnover_community <- turnover_rates %>%
  group_by(site, scale, season, boot_id) %>%
  summarise(
    mean_turnover  = exp(weighted.mean(log(turnover_rate), w = bio_g_m)),
    n_consumers    = n(),
    .groups = "drop"
  )

# -----------------------------------------------------------------------------
# STEP 5: Join with stability metrics
# Align sentinel "annual" back to NA_character_ for season after join
# -----------------------------------------------------------------------------

turnover_stability <- turnover_community %>%
  left_join(
    stability_all %>%
      mutate(season = coalesce(season, "annual")) %>%   # match sentinel
      select(site, scale, season, boot_id, resilience, reactivity, is_stable),
    by = c("site", "scale", "season", "boot_id")
  ) %>%
  mutate(
    season = if_else(season == "annual", NA_character_, season),
    site   = factor(site, levels = c("st14", "oh2", "st7", "st9", "st6", "hver")),
    log_turnover   = log(mean_turnover),
    log_reactivity = log(reactivity)
  )

# Quick check: should have ~n_boots rows per site x scale combination
turnover_stability %>%
  count(site, scale, season)

# =============================================================================
# MODELS
# Site as random intercept — accounts for non-independence within sites
# and lets us ask whether the within-site turnover signal predicts stability
# across the gradient
# =============================================================================

library(lme4)
library(lmerTest)

# Does turnover predict resilience?
m_resilience <- lmer(resilience ~ log_turnover + (1 | site),
                     data = turnover_stability)
summary(m_resilience)

# Does turnover predict reactivity?
m_reactivity <- lmer(log_reactivity ~ log_turnover + (1 | site),
                     data = turnover_stability)
summary(m_reactivity)

# Does the turnover-resilience slope shift across seasons?
# This is the formal test of your drought/pulse-recovery intuition:
# if the slope is steeper in one season, faster-turnover communities
# gain proportionally more resilience during that period
m_seasonal <- lmer(resilience ~ log_turnover * season + (1 | site),
                   data = turnover_stability %>% filter(scale == "seasonal"))
summary(m_seasonal)

# =============================================================================
# PLOTS
# =============================================================================

site_colors <- c(
  st14 = "#2C7BB6", oh2 = "#5AAE61", st7 = "#A6D96A",
  st9  = "#FDAE61", st6 = "#F46D43", hver = "#D73027"
)

season_labels <- c(first = "Jan–Apr", second = "May–Aug", third = "Sep–Dec")

# --- Plot 1: Turnover vs Resilience ---
# Each point is one bootstrap. Black line is the across-gradient trend.
# Site-coloured points show whether the relationship is driven by
# within-site variation or between-site differences in turnover regime.

p_turnover_resilience <- ggplot(
  turnover_stability,
  aes(x = log_turnover, y = resilience, colour = site)
) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 1, se = TRUE) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~scale, labeller = labeller(
    scale = c(annual = "Annual", seasonal = "Seasonal")
  )) +
  labs(
    x     = expression(ln(Turnover~rate)~(day^{-1})),
    y     = expression(lambda[J]^max~"(resilience)"),
    colour = "Site",
    title = "Community turnover rate vs. resilience",
    subtitle = "More negative resilience = faster return after perturbation"
  ) +
  theme_classic(base_size = 13)

# --- Plot 2: Turnover vs Reactivity ---
# Tests whether faster-turnover communities also amplify disturbances more
# before recovering — the opposing force to resilience

p_turnover_reactivity <- ggplot(
  turnover_stability,
  aes(x = log_turnover, y = log_reactivity, colour = site)
) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 1, se = TRUE) +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~scale, labeller = labeller(
    scale = c(annual = "Annual", seasonal = "Seasonal")
  )) +
  labs(
    x     = expression(ln(Turnover~rate)~(day^{-1})),
    y     = expression(log(lambda[H]^max)~"(reactivity)"),
    colour = "Site",
    title = "Community turnover rate vs. reactivity",
    subtitle = "Higher reactivity = larger transient amplification before return"
  ) +
  theme_classic(base_size = 13)

# --- Plot 3: Seasonal slopes ---
# Core test of the drought-recovery question.
# If the slope of turnover on resilience steepens in dry/low season,
# fast-turnover communities gain disproportionate recovery advantage then.

p_seasonal_slopes <- ggplot(
  turnover_stability %>% filter(scale == "seasonal"),
  aes(x = log_turnover, y = resilience, colour = site)
) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 1, se = TRUE) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~season, labeller = labeller(season = season_labels)) +
  labs(
    x      = expression(ln(Turnover~rate)~(day^{-1})),
    y      = expression(lambda[J]^max~"(resilience)"),
    colour = "Site",
    title  = "Does the turnover–resilience slope shift seasonally?",
    subtitle = "A steeper slope in one season = turnover confers more recovery advantage then"
  ) +
  theme_classic(base_size = 13)

p_turnover_resilience
p_turnover_reactivity
p_seasonal_slopes

# =============================================================================
# OPTION 1: Metabolic rate as pace-of-life axis
# Mechanistically cleaner — metrate_J_day_gram IS the Jacobian diagonal
# =============================================================================

# Seasonal: biomass-weighted mean metrate within each season
metrate_seasonal <- community %>%
  filter(!taxon_id %in% basal_taxa, bio_g_m > 0) %>%
  group_by(site, boot_id, season) %>%
  summarise(
    wmean_metrate = weighted.mean(log(metrate_J_day_gram), w = bio_g_m, na.rm = TRUE),
    total_biomass = sum(bio_g_m, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(scale = "seasonal")

# Annual: biomass-weighted mean metrate across all dates, ignoring season
metrate_annual <- community %>%
  filter(!taxon_id %in% basal_taxa, bio_g_m > 0) %>%
  group_by(site, boot_id) %>%
  summarise(
    wmean_metrate = weighted.mean(log(metrate_J_day_gram), w = bio_g_m, na.rm = TRUE),
    total_biomass = sum(bio_g_m, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(scale = "annual", season = "annual")

metrate_community <- bind_rows(metrate_annual, metrate_seasonal)

# Join with stability — coalesce NA season to "annual" sentinel only on
# the stability side, since metrate_community already has explicit labels
metrate_stability <- metrate_community %>%
  left_join(
    stability_all %>%
      mutate(season = coalesce(season, "annual")) %>%
      select(site, scale, season, boot_id, resilience, reactivity),
    by = c("site", "scale", "season", "boot_id")
  ) %>%
  mutate(
    site           = factor(site, levels = c("st14", "oh2", "st7", "st9", "st6", "hver")),
    outlier        = resilience < quantile(resilience, 0.01, na.rm = TRUE),
    log_reactivity = log(reactivity)
  )

# =============================================================================
# MODELS — site-level rather than boot-level
# The vertical clustering shows within-site turnover variance is negligible;
# the signal is between sites. So model at the site level using boot
# distributions to construct uncertainty, rather than pooling all boots.
# =============================================================================

# Summarise to site x season means with bootstrap CIs
stability_metrate_summary <- metrate_stability %>%
  filter(!outlier) %>%                          # remove numerical artifacts
  group_by(site, scale, season) %>%
  summarise(
    metrate_mean     = mean(wmean_metrate,   na.rm = TRUE),
    resilience_mean  = mean(resilience,      na.rm = TRUE),
    resilience_lo    = quantile(resilience, 0.025, na.rm = TRUE),
    resilience_hi    = quantile(resilience, 0.975, na.rm = TRUE),
    reactivity_mean  = mean(log_reactivity,  na.rm = TRUE),
    reactivity_lo    = quantile(log_reactivity, 0.025, na.rm = TRUE),
    reactivity_hi    = quantile(log_reactivity, 0.975, na.rm = TRUE),
    biomass_mean     = mean(total_biomass,   na.rm = TRUE),
    .groups = "drop"
  )

# Linear model across sites — only 6 points per season so treat carefully
# This is a pattern description, not a causal model
m_metrate_resilience <- lm(
  resilience_mean ~ metrate_mean,
  data = stability_metrate_summary %>% filter(scale == "annual")
)
summary(m_metrate_resilience)

m_metrate_seasonal <- lm(
  resilience_mean ~ metrate_mean * season,
  data = stability_metrate_summary %>% filter(scale == "seasonal")
)
summary(m_metrate_seasonal)

# =============================================================================
# PLOTS — site summaries with bootstrap CIs
# Much more readable than plotting all boots; shows real between-site signal
# =============================================================================

site_colors <- c(
  st14 = "#2C7BB6", oh2 = "#5AAE61", st7 = "#A6D96A",
  st9  = "#FDAE61", st6 = "#F46D43", hver = "#D73027"
)

season_labels <- c(
  annual = "Annual",
  first  = "Jan–Apr",
  second = "May–Aug",
  third  = "Sep–Dec"
)

# --- Plot 1: Metabolic rate vs Resilience (annual) ---
p_met_resilience_ann <- stability_metrate_summary %>%
  filter(scale == "annual") %>%
  ggplot(aes(x = metrate_mean, y = resilience_mean, colour = site)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 0.8, se = TRUE) +
  geom_pointrange(aes(ymin = resilience_lo, ymax = resilience_hi), size = 0.8) +
  scale_colour_manual(values = site_colors) +
  labs(
    x      = expression(ln(Mass-specific~metabolic~rate)~(J~g^{-1}~day^{-1})),
    y      = expression(lambda[J]^max~"(resilience)"),
    colour = "Site",
    title  = "Metabolic pace vs. resilience (annual)",
    subtitle = "Points = site means; bars = 95% bootstrap CI"
  ) +
  theme_classic(base_size = 13)

# --- Plot 2: Metabolic rate vs Resilience (seasonal, three panels) ---
p_met_resilience_seas <- stability_metrate_summary %>%
  filter(scale == "seasonal") %>%
  mutate(season = factor(season,
                         levels = c("first", "second", "third"),
                         labels = c("Jan–Apr", "May–Aug", "Sep–Dec"))) %>%
  ggplot(aes(x = metrate_mean, y = resilience_mean, colour = site)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 0.8, se = TRUE) +
  geom_pointrange(aes(ymin = resilience_lo, ymax = resilience_hi), size = 0.7) +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~season) +
  labs(
    x      = expression(ln(Mass-specific~metabolic~rate)~(J~g^{-1}~day^{-1})),
    y      = expression(lambda[J]^max~"(resilience)"),
    colour = "Site",
    title  = "Does the metabolic-rate–resilience slope shift seasonally?",
    subtitle = "A steeper slope = stronger metabolic pace advantage in that season"
  ) +
  theme_classic(base_size = 13)

# --- Plot 3: Metabolic rate vs Reactivity ---
p_met_reactivity <- stability_metrate_summary %>%
  filter(scale == "seasonal") %>%
  mutate(season = factor(season,
                         levels = c("first", "second", "third"),
                         labels = c("Jan–Apr", "May–Aug", "Sep–Dec"))) %>%
  ggplot(aes(x = metrate_mean, y = reactivity_mean, colour = site)) +
  geom_smooth(aes(group = 1), method = "lm",
              colour = "black", linewidth = 0.8, se = TRUE) +
  geom_pointrange(aes(ymin = reactivity_lo, ymax = reactivity_hi), size = 0.7) +
  scale_colour_manual(values = site_colors) +
  facet_wrap(~season) +
  labs(
    x      = expression(ln(Mass-specific~metabolic~rate)~(J~g^{-1}~day^{-1})),
    y      = expression(log(lambda[H]^max)~"(reactivity)"),
    colour = "Site",
    title  = "Metabolic pace vs. reactivity",
    subtitle = "Tests whether faster metabolic pace also amplifies disturbances"
  ) +
  theme_classic(base_size = 13)

p_met_resilience_ann
p_met_resilience_seas
p_met_reactivity
```


```{r}
# =============================================================================
# SUMMARY TABLES
# =============================================================================

# Stability: mean + 95% CI per site x scale x season
stability_summary <- stability_all %>%
  group_by(site, scale, season) %>%
  summarise(
    resilience_mean  = mean(resilience,              na.rm = TRUE),
    resilience_lo    = quantile(resilience, 0.025,   na.rm = TRUE),
    resilience_hi    = quantile(resilience, 0.975,   na.rm = TRUE),
    reactivity_mean  = mean(reactivity,              na.rm = TRUE),
    reactivity_lo    = quantile(reactivity, 0.025,   na.rm = TRUE),
    reactivity_hi    = quantile(reactivity, 0.975,   na.rm = TRUE),
    prop_stable      = mean(is_stable,               na.rm = TRUE),
    n_boots          = n(),
    .groups = "drop"
  )

# Interaction strengths: distribution summary per site x scale x season
interaction_summary <- interactions_all %>%
  group_by(site, scale, season, direction) %>%
  summarise(
    mean_abs_strength = mean(abs(interaction_strength), na.rm = TRUE),
    sd_abs_strength   = sd(abs(interaction_strength),   na.rm = TRUE),
    median_strength   = median(interaction_strength,    na.rm = TRUE),
    n_links           = n(),
    .groups = "drop"
  )

# Proportion of strong interactions per boot (Bazin threshold: top 0.2% of total flux)
# Adapt threshold as needed for your system
strong_flux_threshold <- 0.002  # 0.2% of total flux

strong_flux_summary <- map_dfr(sites, function(s) {
  map_dfr(c("annual", seasons), function(sc) {
    
    fl <- if (sc == "annual") {
      results[[s]]$annual$flux_list
    } else {
      results[[s]]$seasonal[[sc]]$flux_list
    }
    if (is.null(fl)) return(NULL)
    
    map_dfr(names(fl), function(b) {
      mat        <- fl[[b]]
      total_flux <- sum(mat)
      n_strong   <- sum(mat > strong_flux_threshold * total_flux)
      n_total    <- sum(mat > 0)
      tibble(
        site            = s,
        scale           = ifelse(sc == "annual", "annual", "seasonal"),
        season          = ifelse(sc == "annual", NA_character_, sc),
        boot_id         = b,
        total_flux_J    = total_flux,
        n_links         = n_total,
        n_strong_links  = n_strong,
        prop_strong     = n_strong / n_total
      )
    })
  })
})

# Quick checks
glimpse(stability_summary)
glimpse(interaction_summary)
glimpse(strong_flux_summary)

# Access raw objects:
# results[["hver"]]$annual$flux_list             # raw flux matrices
# results[["hver"]]$annual$jacobians             # raw Jacobian matrices
# results[["hver"]]$seasonal$second$stability_summary
# results[["st6"]]$seasonal$first$interaction_strengths

```


```{r}
library(ggplot2)
library(dplyr)
library(forcats)

# ---- helper: ordered label for x-axis ----
stability_summary <- stability_summary %>%
  mutate(
    time_label = case_when(
      scale == "annual"             ~ "Annual",
      season == "first"             ~ "Jan–Apr",
      season == "second"            ~ "May–Aug",
      season == "third"             ~ "Sep–Dec"
    ),
    time_label = factor(time_label, 
                        levels = c("Annual", "Jan–Apr", "May–Aug", "Sep–Dec"))
  )

# ---- Plot 1: Resilience (return rate, more negative = more stable) ----
p_resilience <- ggplot(stability_summary,
                       aes(x = time_label, y = resilience_mean, colour = site)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = resilience_lo, ymax = resilience_hi),
                  position = position_dodge(width = 0.5), size = 0.6) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = NULL, y = expression(lambda[J]^max ~ "(resilience)"),
    title = "Resilience across sites and temporal scales",
    subtitle = "More negative = faster return to equilibrium after perturbation",
    colour = "Site"
  ) +
  theme_classic(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# ---- Plot 2: Reactivity (log scale due to magnitude) ----
p_reactivity <- ggplot(stability_summary,
                       aes(x = time_label, y = reactivity_mean, colour = site)) +
  geom_pointrange(aes(ymin = reactivity_lo, ymax = reactivity_hi),
                  position = position_dodge(width = 0.5), size = 0.6) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = NULL, y = expression(lambda[H]^max ~ "(reactivity, log scale)"),
    title = "Reactivity across sites and temporal scales",
    subtitle = "Higher = larger transient amplification before return to equilibrium",
    colour = "Site"
  ) +
  theme_classic(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# ---- Plot 3: Resilience vs Reactivity scatter (stability space) ----
p_space <- ggplot(stability_summary,
                  aes(x = resilience_mean, y = log10(reactivity_mean),
                      colour = site, shape = scale, label = time_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(size = 3) +
  ggrepel::geom_text_repel(size = 3, show.legend = FALSE) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = expression(lambda[J]^max ~ "(resilience)"),
    y = expression(log[10](lambda[H]^max) ~ "(reactivity)"),
    title = "Stability space",
    subtitle = "Bottom-left = stable AND non-reactive (most stable)",
    colour = "Site", shape = "Scale"
  ) +
  theme_classic(base_size = 13)

p_resilience
p_reactivity
p_space

```




```{r}

library(tidyverse)

site_colors <- c(
  st14 = "#2C7BB6",
  oh2  = "#5AAE61",
  st7  = "#A6D96A",
  st9  = "#FDAE61",
  st6  = "#F46D43",
  hver = "#D73027"
)

# Summarise: total community biomass per site x date x boot
# (sum across taxa within each bootstrap replicate, then summarise boots)
community_biomass <- community %>%
  filter(!taxon_id %in% basal_taxa) %>%        # exclude placeholder basal rows
  group_by(site, date_id, boot_id) %>%
  summarise(total_bio_g_m = sum(bio_g_m, na.rm = TRUE), .groups = "drop") %>%
  group_by(site, date_id) %>%
  summarise(
    mean_bio  = mean(total_bio_g_m, na.rm = TRUE),
    lo_bio    = quantile(total_bio_g_m, 0.025, na.rm = TRUE),
    hi_bio    = quantile(total_bio_g_m, 0.975, na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  mutate(
    doy  = yday(date_id),                       # day of year strips the actual year
    site = factor(site,
                  levels = c("st14", "oh2", "st7", "st9", "st6", "hver"))
  )

# Month breaks for x-axis
month_breaks <- yday(as.Date(paste0("2011-", 1:12, "-01")))
month_labels <- month.abb

p <- ggplot(community_biomass, aes(x = doy, y = mean_bio, colour = site, fill = site)) +
  geom_ribbon(aes(ymin = lo_bio, ymax = hi_bio), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  scale_colour_manual(values = site_colors) +
  scale_fill_manual(values = site_colors) +
  scale_x_continuous(
    breaks = month_breaks,
    labels = month_labels,
    limits = c(1, 365)
  ) +
  scale_y_log10(labels = scales::label_comma()) +
  labs(
    x      = NULL,
    y      = expression("Total community biomass (g m"^{-2}*")"),
    colour = "Site (cold → warm)",
    fill   = "Site (cold → warm)",
    caption = "Shaded bands = 95% bootstrap CI"
  ) +
  theme_classic(base_size = 13) +
  theme(
    legend.position   = "none",
    axis.text.x       = element_text(size = 10),
    plot.caption      = element_text(size = 9, colour = "grey50", hjust = 0)
  )

p

# Save tiff
ggsave(
  filename    = "figures/seasonal_bio_nolegend.tiff",
  plot        = p,
  width       = 8,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"
)

```


### van Althena et al. 2015 - even distributions of interactions confer stability 
G = 0: all links carry equal flux / equal interaction strength
G = 1: one link carries all the flux / one interaction dominates completely

van Althena major Results:
G_F ranges 0.86–0.97 across 15 real soil food webs — all highly skewed, few strong fluxes dominate
G_F and G_I are positively correlated (ρ = 0.58) — skewed fluxes produce skewed interactions
Both G_F and G_I are positively correlated with instability (ρ = 0.78 and 0.60) — more skewed distributions → less stable webs

```{r}

# Gini function (Sen 1973 definition, matches van Altena eq. 7-8)
gini <- function(x) {
  x <- abs(x[is.finite(x) & x != 0])
  if (length(x) < 2) return(NA_real_)
  n  <- length(x)
  mu <- mean(x)
  sum(abs(outer(x, x, "-"))) / (2 * n^2 * mu)
}

# G_F: Gini of flux distributions, per boot
# 
gini_flux <- map_dfr(sites, function(s) {
  map_dfr(c("annual", seasons), function(sc) {
    
    fl <- if (sc == "annual") results[[s]]$annual$flux_list
          else results[[s]]$seasonal[[sc]]$flux_list
    if (is.null(fl)) return(NULL)
    
    map_dfr(names(fl), function(b) {
      tibble(
        site    = s,
        scale   = ifelse(sc == "annual", "annual", "seasonal"),
        season  = ifelse(sc == "annual", NA_character_, sc),
        boot_id = b,
        G_F     = gini(as.vector(fl[[b]]))
      )
    })
  })
})

# G_I: Gini of off-diagonal |J_ij|, but only consumer-resource interactions
gini_interactions <- interactions_real %>%
  group_by(site, scale, season, boot_id) %>%
  summarise(G_I = gini(interaction_strength), .groups = "drop")

# Join and summarise
gini_summary <- left_join(
  gini_flux,
  gini_interactions,
  by = c("site", "scale", "season", "boot_id")
) %>%
  group_by(site, scale, season) %>%
  summarise(
    G_F_mean = mean(G_F, na.rm = TRUE),
    G_F_lo   = quantile(G_F, 0.025, na.rm = TRUE),
    G_F_hi   = quantile(G_F, 0.975, na.rm = TRUE),
    G_I_mean = mean(G_I, na.rm = TRUE),
    G_I_lo   = quantile(G_I, 0.025, na.rm = TRUE),
    G_I_hi   = quantile(G_I, 0.975, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  mutate(
    site = factor(site, levels = c("st14","oh2","st7","st9","st6","hver")),
    time_label = case_when(
      scale == "annual" ~ "Annual",
      season == "first" ~ "Jan–Apr",
      season == "second" ~ "May–Aug",
      season == "third" ~ "Sep–Dec"
    ) %>% factor(levels = c("Annual","Jan–Apr","May–Aug","Sep–Dec"))
  )
```
Results: G_F range from 0.81-0.94, G_I range from 0.63-0.83.

# Visualizing the skew of fluxes and interaction strengths 

```{r}
site_order  <- c("st14","oh2","st7","st9","st6","hver")
site_colors <- c(st14="#2C7BB6", oh2="#5AAE61", st7="#A6D96A",
                 st9="#FDAE61", st6="#F46D43", hver="#D73027")

# ── Plot 1: Annual G_F and G_I across the thermal gradient ──────────────────
# The primary cross-site comparison — most interpretable panel
# Faceted so G_F and G_I sit side by side on the same scale

gini_annual <- gini_summary %>% filter(scale == "annual")
gini_seasonal <- gini_summary %>% filter(scale == "seasonal")

p_annual <- gini_annual %>%
  select(site, G_F_mean, G_F_lo, G_F_hi, G_I_mean, G_I_lo, G_I_hi) %>%
  pivot_longer(
    cols      = -site,
    names_to  = c("metric", ".value"),
    names_pattern = "(G_F|G_I)_(mean|lo|hi)"
  ) %>%
  mutate(metric = recode(metric,
                         G_F = "GF (flux)",
                         G_I = "GI (interaction strength)")) %>%
  ggplot(aes(x = site, y = mean, color = site)) +
  geom_pointrange(aes(ymin = lo, ymax = hi), size = 0.7) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  facet_wrap(~metric) +
  labs(x = "Site", y = "Gini coefficient",
       title = "Interaction inequality across the thermal gradient",
       subtitle = "G = 0: all links carry equal flux / equal interaction strength
G = 1: one link carries all the flux / one interaction dominates completely",
       color = "Site") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")


p_seasonal <- gini_seasonal %>%
  ggplot(aes(x = time_label, y = G_F_mean, color = site)) +
  geom_pointrange(aes(ymin = G_F_lo, ymax = G_F_hi), size = 0.7) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  labs(x = "Season", y = "Gini coefficient",
       title = "Flux inequality across the thermal gradient",
       subtitle = "G = 0: all links carry equal flux
G = 1: one link carries all the flux",
       color = "Site") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")


pd <- position_dodge(width = 0.4)

# G_F version
p_seasonal_GF <- gini_summary %>%
  filter(scale == "seasonal") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = time_label, y = G_F_mean, 
             color = site, group = site)) +
  geom_line(aes(group = site), position = pd, 
            linewidth = 0.6, alpha = 0.5) +
  geom_pointrange(aes(ymin = G_F_lo, ymax = G_F_hi),
                  position = pd, size = 0.6) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  labs(x = "Season", 
       y = expression(G[F]),
       title = "Flux inequality across seasons",
       subtitle = "G = 0: all links equal  |  G = 1: one link dominates",
       color = "Site") +
  theme_classic(base_size = 13)

# G_I version — identical but swap G_F → G_I
p_seasonal_GI <- gini_summary %>%
  filter(scale == "seasonal") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = time_label, y = G_I_mean,
             color = site, group = site)) +
  geom_line(aes(group = site), position = pd,
            linewidth = 0.6, alpha = 0.5) +
  geom_pointrange(aes(ymin = G_I_lo, ymax = G_I_hi),
                  position = pd, size = 0.6) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  labs(x = "Season",
       y = expression(G[I]),
       title = "Interaction strength inequality across seasons",
       subtitle = "G = 0: all interactions equal  |  G = 1: one interaction dominates",
       color = "Site") +
  theme_classic(base_size = 13)

# Combined — shared legend on right
p_seasonal_GF + p_seasonal_GI + 
  plot_layout(guides = "collect") &
  theme(legend.position = "right")


# p_annual_GF
p_annual_GF <- gini_summary %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = site, y = G_F_mean, color = site)) +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.86, ymax = 0.97,
           alpha = 0.08, fill = "grey30") +
  geom_pointrange(aes(ymin = G_F_lo, ymax = G_F_hi), size = 0.7) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  labs(x = "Site (cold → warm)", 
       y = expression(G[F]),
       title = "Annual flux inequality",
       subtitle = "Grey = van Altena range (0.86–0.97)",
       color = "Site") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

# p_annual_GI
p_annual_GI <- gini_summary %>%
  filter(scale == "annual") %>%
  mutate(site = factor(site, levels = site_order)) %>%
  ggplot(aes(x = site, y = G_I_mean, color = site)) +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.64, ymax = 0.86,
           alpha = 0.08, fill = "grey30") +
  geom_pointrange(aes(ymin = G_I_lo, ymax = G_I_hi), size = 0.7) +
  scale_color_manual(values = site_colors) +
  scale_y_continuous(limits = c(0.6, 1.0), breaks = seq(0.6, 1.0, 0.1)) +
  labs(x = "Site (cold → warm)", 
       y = expression(G[I]),
       title = "Annual interaction strength inequality",
       subtitle = "Grey = van Altena range (0.64–0.86)",
       color = "Site") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

# 4-panel layout
Gini <- (p_seasonal_GF | p_annual_GF) / (p_seasonal_GI | p_annual_GI) +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")


# Save tiff
ggsave(
  filename    = "figures/Gini.tiff",
  plot        = Gini,
  width       = 8,
  height      = 4.5,
  units       = "in",
  dpi         = 300,
  compression = "lzw"  
)

ggsave(
  filename = "figures/Gini.pdf",
  plot     = Gini,
  width    = 12,
  height   = 9,
  units    = "in"
)
```




